---
title: "Visium data - Comparison of clustering oberlap between 'baseline' and 'insulin' analyses"
author: "L. Franzén lovisa.franzen@scilifelab.se"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    theme: cosmo
    highlight: tango
    css: style.css
    code_folding: "hide"
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    number_sections: false
    
---

***

# Initialize  

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = F,
	cache.lazy = FALSE
)

set.seed(42)

#' Load libs

library(Matrix)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(ggrepel)
library(cowplot)
library(patchwork)
library(ggpubr)
library(writexl)
library(Seurat)
library(STutility)

library(DT)
library(knitr)


#' Define analysis tag
ANALYSIS_ID <- "visium_comparison"


#' Define project paths
PROJECT_ID <- "visium"
DIR_WD <- getwd()
DIR_ROOT <- file.path(getwd(), "..")  # 

DIR_DATA <- file.path(DIR_ROOT, "data", PROJECT_ID)
DIR_RES <- file.path(DIR_ROOT, "results" , PROJECT_ID)
DIR_FIG <- file.path(DIR_RES, "figures")


#' Colors
source(file.path(DIR_WD, "colors.R"))

#' Define custom functions
'%!in%' <- function(x,y)!('%in%'(x,y));
```


<br><br> 

## Load necessary files and tables  
  
  
### Sample insulin cluster annotation table  

```{r read_canno}
canno_base <- read.csv(file.path(DIR_RES, "tables", "visium_baseline.clustering_annotations.csv"), stringsAsFactors = F)
canno_ins <- read.csv(file.path(DIR_RES, "tables", "visium_insulin.clustering_annotations.csv"), stringsAsFactors = F)
```
  
<br><br>

### Seurat object  

```{r read_seobjs}
se_base <- readRDS(file.path(DIR_RES, "se-object.visium_baseline.rds"))
se_ins <- readRDS(file.path(DIR_RES, "se-object.visium_insulin.rds"))

#' minor formatting
se_base <- AddMetaData(se_base, metadata = as.factor(as.numeric(as.character(se_base[[]][, "seurat_clusters"]))), col.name = "seurat_clusters_f" )
```

<br><br>

***

# Subset overlap

```{r overlap_subset}
tissue_id_keep <- c("NK50-before-A","NO148-before","NO137-before","NK49-before-B")

se_compare <- subset(se_ins, tissue_id %in% tissue_id_keep)
se_baseline_subset <- subset(se_base, tissue_id %in% tissue_id_keep)

new_names <- data.frame(tissue_id = se_compare[[]]$tissue_id,
                        barcodes_original = rownames(se_compare[[]]),
                        apply(do.call(rbind, strsplit(rownames(se_compare[[]]), split = "_")), 2, as.character),
                        stringsAsFactors = F)
new_names_base <- data.frame(tissue_id = se_baseline_subset[[]]$tissue_id,
                             barcodes_original = rownames(se_baseline_subset[[]]),
                             apply(do.call(rbind, strsplit(rownames(se_baseline_subset[[]]), split = "_")), 2, as.character),
                             stringsAsFactors = F)
new_names$rep = 0
new_names_base$rep = 0
i=1
for(s in tissue_id_keep) {
  new_names[new_names$tissue_id == s, "rep"] <- i
  new_names_base[new_names_base$tissue_id == s, "rep"] <- i
  i = i+1
}
new_names$barcode_new <- paste(new_names$X1, new_names$rep, sep = "_")
rownames(new_names) <- new_names$barcode_new
new_names_base$barcode_new <- paste(new_names_base$X1, new_names_base$rep, sep = "_") 
rownames(new_names_base) <- new_names_base$barcode_new
new_names_base <- new_names_base[intersect(rownames(new_names), rownames(new_names_base)), ]

se_baseline_subset <- SubsetSTData(se_baseline_subset, spots = new_names_base$barcodes_original)
metadata_baseline <- se_baseline_subset[[]][new_names_base$barcodes_original, ]

metadata_baseline$barcodes_original <- new_names_base$barcodes_original
rownames(metadata_baseline) <- as.character(new_names_base$barcode_new)

metadata_insulin <- se_compare[[]][new_names$barcodes_original, ]
rownames(metadata_insulin) <- new_names$barcode_new
```


Number of spots in overlapping data:
```{r view_metadata_dim}
dim(metadata_baseline)
dim(metadata_insulin)
```


Overlapping samples:
```{r check_subset}
unique(se_compare$subject_id)
unique(se_compare$tissue_id)

unique(se_compare$sample_id)
unique(se_compare$novaseq_id)
```

```{r prep_overlap_df}
row_overlap <- intersect(rownames(metadata_baseline), rownames(metadata_insulin))
metadata_baseline <- metadata_baseline[row_overlap, ]
metadata_insulin <- metadata_insulin[row_overlap, ]

cluster_res <- "seurat_clusters" # "SCT_snn_res.0.7"

cluster_comparison <- data.frame(cluster_baseline = as.factor(as.numeric(as.character(metadata_baseline$seurat_clusters))),
                                 cluster_insulin = as.factor(as.numeric(as.character(metadata_insulin[, cluster_res]))),
                                 metadata_baseline[, colnames(metadata_baseline)[c(2:23, 36, 45:48)]], 
                                 stringsAsFactors = F)

# se_compare <- AddMetaData(se_compare, metadata = metadata_baseline$seurat_clusters_f, col.name = "cluster_baseline")
```
<br><br>

# Quantify overlap

## All four subjects pooled  

```{r compare_clusters}
d_plot <- cluster_comparison %>%
  group_by(cluster_insulin, cluster_baseline) %>%  #tissue_id, 
  dplyr::count()

datatable(d_plot, rownames = F)
```

```{r compare_clusters_plot_pct, fig.width=12, fig.height=6}
p1 <- ggplot(d_plot, 
             aes(x = cluster_insulin, y = n, fill = cluster_baseline)) +
  geom_bar(stat = 'identity', color=NA, position = "fill") + # position:"fill" (pct) or "stack" (count)
  scale_fill_manual(values = canno_base[order(canno_base$seurat_clusters), "cluster_color"]) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  legend.position = "top")


p2 <- ggplot(d_plot, 
             aes(x = cluster_baseline, y = n, fill = cluster_insulin)) +
  geom_bar(stat = 'identity', color=NA, position = "fill") + # position:"fill" (pct) or "stack" (count)
  scale_fill_manual(values = canno_ins[order(canno_ins$seurat_clusters), "cluster_color"]) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top")

p1 + p2
```
<br><br>

## Per subject  

```{r compare_clusters_subject}
d_plot_sub <- cluster_comparison %>%
  group_by(cluster_insulin, cluster_baseline, subject_id) %>%  #tissue_id, 
  dplyr::count()

datatable(d_plot_sub, rownames = F)
```

<br><br>


Percentage:  
```{r compare_clusters_subject_plot_pct, fig.width=12, fig.height=8}
p1 <- ggplot(d_plot_sub, 
             aes(x = cluster_insulin, y = n, fill = cluster_baseline)) +
  geom_bar(stat = 'identity', color=NA, position = "fill") + # position:"fill" (pct) or "stack" (count)
  scale_fill_manual(values = canno_base[order(canno_base$seurat_clusters), "cluster_color"]) +
  facet_wrap(~subject_id) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  legend.position = "top");p1


p2 <- ggplot(d_plot_sub, 
             aes(x = cluster_baseline, y = n, fill = cluster_insulin)) +
  geom_bar(stat = 'identity', color=NA, position = "fill") + # position:"fill" (pct) or "stack" (count)
  scale_fill_manual(values = canno_ins[order(canno_ins$seurat_clusters), "cluster_color"]) +
  facet_wrap(~subject_id) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top");p2
```
<br><br>


Raw spot count:  
```{r compare_clusters_subject_plot_n, fig.width=12, fig.height=8}
p1 <- ggplot(d_plot_sub, 
             aes(x = cluster_insulin, y = n, fill = cluster_baseline)) +
  geom_bar(stat = 'identity', color=NA, position = "stack") + # position:"fill" (pct) or "stack" (count)
  scale_fill_manual(values = canno_base[order(canno_base$seurat_clusters), "cluster_color"]) +
  scale_y_log10() +
  facet_wrap(~subject_id) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  legend.position = "top");p1


p2 <- ggplot(d_plot_sub, 
             aes(x = cluster_baseline, y = n, fill = cluster_insulin)) +
  geom_bar(stat = 'identity', color=NA, position = "stack") + # position:"fill" (pct) or "stack" (count)
  scale_fill_manual(values = canno_ins[order(canno_ins$seurat_clusters), "cluster_color"]) +
  scale_y_log10() +
  facet_wrap(~subject_id) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top");p2
```
<br><br>


## Export tables

```{r export_tables}
write.csv(d_plot, file = file.path(DIR_RES, "tables", "visium_comparison.baseline_insulin_cluster_overlap.csv"), row.names = F)
write.csv(d_plot_sub, file = file.path(DIR_RES, "tables", "visium_comparison.baseline_insulin_cluster_overlap_subject.csv"), row.names = F)
```
<br><br>


## Cluster correlation

```{r cluster_corr}
# library(corrplot)
# library(ggcorrplot)
library(pheatmap)

clus_ins <- as.numeric(unique(d_plot$cluster_insulin))
clus_base <- as.numeric(unique(d_plot$cluster_baseline))
corr_df <- matrix(0, nrow = length(clus_ins), ncol = length(clus_base))

for(ci in clus_ins) {
  # print(paste("row:", r))
  for(cb in clus_base) {
    # print(paste("column:", c))
    tmp <- subset(d_plot, cluster_insulin == as.character(ci) & cluster_baseline == as.character(cb))
    # print(dim(tmp))
    if(dim(tmp)[1] != 0) {
      corr_df[ci, cb] <- tmp$n
    }
  }
}
```

```{r cluster_corr_plot, fig.width=4.8, fig.height=4.4}
rownames(corr_df) <- paste0("i_C", clus_ins)
colnames(corr_df) <- paste0("b_C", clus_base)

corr_df_pct <- corr_df
for(r in 1:nrow(corr_df)) {
  corr_df_pct[r, ] <-  round(corr_df[r, ] / (colSums(corr_df)+0.01), 2)
}

ph_col <- c("#ffffff", colorRampPalette(brewer.pal(9, "OrRd"))(20))
pheatmap(corr_df_pct, 
         color = ph_col, 
         border_color = "grey80", 
         cluster_cols = F,
         cluster_rows = F,
         main = "Cluster overlap (% of baseline spots)")
```

<br><br>

***  

# Wrap up

<br><br>

## Session information  
  
This analysis was last compiled on `r Sys.Date()`.  
  
<br>
  
```{r sessioninfo}
sessionInfo()
```

