---
title: "Visium pseudobulk analysis and comparison against CAGE data"
author: "L. Franzén lovisa.franzen@scilifelab.se"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    theme: cosmo
    highlight: tango
    css: style.css
    code_folding: "hide"
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    number_sections: false
---

  
***
  
**Description**  

Psudo-bulk analysis of Visium data by summing gene expression data for all spots within a section/subject.  

TODO: Add info about CAGE data. CAGE: DE analysis between before/after insulin clamp from abdominal scWAT tissue (?).


***
  
  
# Set up  
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = T,
	cache.lazy = FALSE
)

set.seed(42)

#' Load libs
# BiocManager::install("apeglm")

library(Matrix)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(ggrepel)
library(cowplot)
library(patchwork)
library(pheatmap)
library(ggpubr)
library(writexl)

library(Seurat)
library(STutility)

library(DT)
library(knitr)

library(DESeq2)
library(biomaRt)


#' Define project paths
PROJECT_ID <- "pseudobulk_cage"
DIR_WD <- getwd()
DIR_ROOT <- file.path(getwd(), "..")  # 

DIR_DATA <- file.path(DIR_ROOT, "data")
DIR_RES <- file.path(DIR_ROOT, "results" , PROJECT_ID)
DIR_FIG <- file.path(DIR_RES, "figures")


#' Colors
source(file.path(DIR_WD, "colors.R"))

#' Define custom functions
'%!in%' <- function(x,y)!('%in%'(x,y))
```
  

   
<br><br>
  
## Read CAGE data
  
```{r read_cage}
cage <- read.table(file.path(DIR_ROOT, "data", "CAGE", "DE_bvsa_all_FDR5.txt"), header = T, sep = "\t", stringsAsFactors = F)
cage_sign <-subset(cage, FDR<0.05)
cage_sign$chr_loc <- rownames(cage_sign)
cage_sign <- cage_sign[-grep(pattern = "None@", x = cage_sign$Annotation), ]
```

```{r cage_biomart}
cage_gene_ann <- as.character(cage_sign$Annotation)

cage_ann_data <- data.frame(do.call(rbind, lapply(strsplit(cage_gene_ann, split = "@"), function(x) head(x, n=2))), stringsAsFactors = F)
cage_ann_data[, 3] <- data.frame(do.call(rbind, lapply(strsplit(cage_ann_data$X2, split = ":"), function(x) head(x, n=2))), stringsAsFactors = F)
colnames(cage_ann_data) <- c("gene_symbol", "annotation2", "enst_v_id")
cage_ann_data$Annotation <- cage_gene_ann
cage_ann_data$enst_id <- gsub(pattern = "\\.[0-9]*", replacement = "", x = cage_ann_data$enst_v_id)
cage_ann_data <- cage_ann_data[cage_ann_data$enst_v_id != "None", ]

cage_enst_ids <- unique(cage_ann_data$enst_id)
ensembl <- useMart("ensembl",dataset="hsapiens_gene_ensembl")
bm <- getBM(attributes = c("ensembl_transcript_id_version", "ensembl_transcript_id", "hgnc_symbol", "gene_biotype"),
            filters = "ensembl_transcript_id", 
            values = cage_enst_ids,
            mart = ensembl
            )

cage_ann_data2 <- merge(x = cage_ann_data, y = bm, by.x = "enst_id", by.y = "ensembl_transcript_id", all=T)

cage_sign <- merge(x = cage_sign, y = cage_ann_data2, by.x = "Annotation")
dim(cage_sign)
```


## Read visium se objects

```{r read_visium_se}
DIR_VISIUM_RES <- file.path(DIR_ROOT, "results", "visium")
se_base <- readRDS(file.path(DIR_VISIUM_RES, "se-object.visium_baseline.rds"))
se_ins <- readRDS(file.path(DIR_VISIUM_RES, "se-object.visium_insulin.rds"))

canno_base <- read.csv(file.path(DIR_VISIUM_RES, "tables", "visium_baseline.clustering_annotations.csv"), stringsAsFactors = F)
canno_ins <- read.csv(file.path(DIR_VISIUM_RES, "tables", "visium_insulin.clustering_annotations.csv"), stringsAsFactors = F)
```
  
  
## Read visium sample meta data and gene information  

### Meta data 

Based on information about the samples, a file called "sample_metadata.tsv" was created to contain all necessary information about the samples.  
  
```{r read_metadata}
metadata <- read.table(file = file.path(DIR_DATA, "visium", "visium_sample_metadata.tsv"), sep = "\t", header = T, stringsAsFactors = F)
metadata$bmi <- as.numeric(metadata$bmi)
rownames(metadata) <- metadata$sample_name

datatable(metadata, rownames = F, options = list(scrollX = TRUE), caption = "Sample Metadata")
```
<br><br>
  
  
### Gene information

Gene information is obtained from a separate table containing the human genome with ensembl ids, hgnc symbols, and gene biotype.  

```{r read_genes} 
gene_table <- as.data.frame( read.delim(file = file.path(DIR_DATA, "gene_annotation", "genes.tsv"), stringsAsFactors = F))

gene_table$gene_id <- gsub(pattern = "\\.[0-9]*", replacement = "", x = gene_table$gene_id)
rownames(gene_table) <- gene_table$gene_id

head(gene_table)
```

<br><br>

  
# Bulk analysis
  
## Visium Pseudo-bulk: Pool data  

To get the bulk data, the gene expression across all spots within a sample will be summed.  
  
### Baseline visium data  

```{r pool_data_per_sample_base}
metadata_base <- subset(metadata, insulin_stim == 0 | is.na(insulin_stim))

bulk_data_base <- data.frame(row.names = rownames(se_base))
s_names <- metadata_base$sample_name

for (s in s_names ) {
  s_spots <- colnames(subset(se_base, sample_name == s))
  st_data_s <- SubsetSTData(se_base, spots = s_spots)
  st_data_s_mat <- st_data_s@assays$RNA@counts
  s_data <- data.frame(Matrix::rowSums(st_data_s_mat))
  colnames(s_data) <- s
  
  bulk_data_base <- cbind(bulk_data_base, s_data)
}
head(bulk_data_base)
```
  
```{r save_bulk_data_base}
write.table(x = bulk_data_base, file = file.path(DIR_RES, "tables", "pseudobulk_visium.baseline.tsv"), quote = F, sep = "\t", row.names = T)
```
  
<br>
  
### Insulin visium data
  
```{r pool_data_per_sample_ins}
metadata_ins <- subset(metadata, insulin_stim == 0 | insulin_stim == 1)
metadata_ins$subject_alias_ins <- paste0(metadata_ins$subject_alias, "_", metadata_ins$insulin_stim)

bulk_data_ins <- data.frame(row.names = rownames(se_ins))
s_names <- metadata_ins$sample_name

for (s in s_names ) {
  s_spots <- colnames(subset(se_ins, sample_name == s))
  st_data_s <- SubsetSTData(se_ins, spots = s_spots)
  st_data_s_mat <- st_data_s@assays$RNA@counts
  
  s_data <- data.frame(Matrix::rowSums(st_data_s_mat))
  colnames(s_data) <- s
  
  bulk_data_ins <- cbind(bulk_data_ins, s_data)
}
head(bulk_data_ins)
```
  
```{r save_bulk_data_ins}
write.table(x = bulk_data_ins, file = file.path(DIR_RES, "tables", "pseudobulk_visium.insulin.tsv"), quote = F, sep = "\t", row.names = T)
```
  
  
  
## Differential gene expression

I'll perform the DE analysis using the DESeq2 R package. Vignette available [here](https://bioconductor.riken.jp/packages/3.7/bioc/vignettes/DESeq2/inst/doc/DESeq2.html).  
  
The lean/obese comparison will be made using data from all samples which haven't been subjected to insulin treatment. The male subjects are not included for now, to make the comparison more fair.  
  
The insulin untreated/treated comparison is made using the samples from individuals who have been subjected to insulin treatment and where tissue samples were collected before and afterwards. This data is therefore paired, which will be taken into account in the design formula.  


### Column data  

```{r de_design_oble}
#' Coldata/Design for obese vs lean
de_des_oble <- subset(metadata_base, (bmi < 28.0 | bmi > 30.0), c("condition", "gender", "date_exp", "tissue_id", "subject_id", "subject_alias"))
de_des_oble$date_exp <- gsub(pattern = "-", replacement = ".", x = de_des_oble$date_exp)
de_des_oble$tissue <- de_des_oble$subject_id
de_des_oble

#' Coldata/Design for insulin stimulation
de_des_ins <- metadata_ins[, c("insulin_stim", "tissue_id", "condition", "gender", "date_exp", "subject_id", "subject_alias", "subject_alias_ins")]
de_des_ins$date_exp <- gsub(pattern = "-", replacement = ".", x = de_des_ins$date_exp)
de_des_ins$insulin_stim <- as.factor(de_des_ins$insulin_stim)
de_des_ins$tissue <- unlist( lapply( strsplit(x = de_des_ins$tissue_id, split = "-"), function(x) head(x, n=1) ) )
de_des_ins$tissue_id <- NULL
de_des_ins <- de_des_ins[2:dim(de_des_ins)[1],]  # remove NK49-before from first batch
de_des_ins
```

<br><br>

### Lean vs obese

Performing DEA based on 'condition', with 'date_exp' specified as covariates (add 'gender' if both male and female subjects are included).


#### DESeq2

```{r deseq_oble, message=TRUE}
s_oble <- rownames(de_des_oble)
d_oble <- bulk_data_base[, s_oble]
# rownames(d_oble) <- gene_table[rownames(d_oble), "gene_name_2"]

dds_oble <- DESeqDataSetFromMatrix(countData = d_oble,
                                   colData = de_des_oble,
                                   design = ~ date_exp + condition)  #  + gender

dds_oble <- DESeq(dds_oble)

resultsNames(dds_oble) # lists the coefficients

res_oble <- results(dds_oble, name="condition_obese_vs_lean")
res_oble

resLFC_oble <- lfcShrink(dds_oble, coef="condition_obese_vs_lean", type="apeglm")
resLFC_oble
```
  
<br><br>
  
#### Plot results  

`plotMA()` shows the log2 fold changes attributable to a given variable over the mean of normalized counts for all the samples in the DESeqDataSet. Points will be colored red if the adjusted p value is less than 0.1. Points which fall out of the window are plotted as open triangles pointing either up or down.


```{r deseq_oble_plot2, fig.width=6, fig.height=6, fig.show='hold'}
#' MA-plot
plotMA(res_oble, ylim=c(-2,2), main = "MA-plot")
plotMA(resLFC_oble, main = "MA-plot with shrunken log2 fc (apeglm)")
```

```{r deseq_oble_restable_plot, fig.width=3, fig.height=2, fig.show='hold'}
res_oble_df <- as.data.frame(res_oble)

res_oble_head <- res_oble_df %>% 
  dplyr::arrange(padj, abs(log2FoldChange)) %>%
  dplyr::filter(padj<0.05) %>%
  head(10)
res_oble_head

d_plot <- t(d_oble)
d_plot <- cbind(d_plot, de_des_oble)

for(g in rownames(res_oble_head)){
  p <- ggplot(d_plot, aes_string(x = "condition", y = g)) +
    geom_boxplot() +
    geom_point(aes(color=tissue)) +
    scale_color_manual(values = colors_multi) +
    theme_classic()
  print(p)
}
```
  
```{r deseq_oble_vulcano, fig.width=6, fig.height=5, fig.show='hold'}
#' vulcano plot
res_oble_df$significant_0.05 <- res_oble_df$padj < 0.05 
res_oble_df$gene_labels <- ifelse(res_oble_df$padj < 0.05,
                                  rownames(res_oble_df), "")

glabs_oble <- res_oble_df[res_oble_df$gene_labels != "" & !is.na(res_oble_df$gene_labels), "gene_labels"]

minmax_xilm <- round( max(abs(res_oble_df$log2FoldChange)) + 1 )

p <- ggplot(res_oble_df, aes(x = log2FoldChange, y = -log10(padj), color = significant_0.05)) +
          # geom_hline(yintercept = -log10(0.05), linetype="dashed", color = colors_multi[3], alpha=.8) +
          geom_hline(yintercept = 0, linetype="solid", color = "grey90") +
          geom_vline(xintercept = c(-2, 2), linetype="dashed", color = "grey90") +
          geom_point(size = .5, alpha = 0.8) +
          geom_text_repel(data = res_oble_df, aes(x = log2FoldChange, y = -log10(padj), label = gene_labels), color = colors_multi[3], size = 3) +
          scale_color_manual(values = colors_multi[c(1,3)]) +
          xlim(c(-minmax_xilm, minmax_xilm)) +
          labs(title = "DE genes: lean vs obese") +
          theme_classic() +
          theme(legend.position = "none",
                plot.title = element_text(size = rel(1.5), hjust = 0.5),
                axis.title = element_text(size = rel(1.25)),
                axis.line = element_line(colour = "grey50"));p

fname <- "bulk_dea_oble_vulcano"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 6, height = 5, useDingbats = F);p;dev.off()
tiff(filename = file.path(DIR_FIG, paste0(fname, ".tiff")), units = "cm", width = 6*2.5, height = 5*2.5, res = 600); p; dev.off()
```

  
*Vulcano plot of differentially expressed genes when comparing obese vs lean individuals. The horizontal yellow dashed line corresponds to -log10(0.05), and the vertical grey dashed lines correspond to log2FC = 2.*
  
  
#### Results table
  
View and export results table
  
```{r deseq_oble_res_table}
#' table with exp values for genes labelled in above vulcano plot
de_des_oble[, c("tissue_id", "condition")]
# datatable(d_oble[glabs_oble, rownames(de_des_oble)], rownames = T, options = list(scrollX = TRUE), caption = "Expression of labelled genes in above vulcano plot.")
```
  

```{r deseq_oble_res_export}
table_export <- res_oble_df[order(res_oble_df$padj), ]
table_export$gene <- rownames(table_export)
table_export <- table_export[, c("gene", colnames(table_export)[-7])]


datatable(subset(table_export, padj<0.05), rownames = F, options = list(scrollX = TRUE), caption = "Bulk DEA: Lean vs Obese. Significant genes (padj<0.05)")


fname <- "bulk_dea_oble.tsv"
write.table(x = table_export, file = file.path(DIR_RES, "tables", fname), quote = F, sep = "\t")
```
  
<br><br>
  
  
  
### Insulin: Untreated vs treated

Performing DEA based on 'insulin_stim', with 'tissue' specified as covariate in order for it to take the paired data into account. 
'Condition' is also included as a factor in order to see how insulin stimulation depends on condition (obese or lean).  

The design matrix in this case becomes a little bit more complicated. We can look at both the effect of insulin treatment in lean or obese separately, but we can also combine it and look at how insulin treatment differs between lean and obese individuals. Which approach to take depends on the question we would like to answer here.  
  
  
#### DESeq2
  
```{r deseq_ins, message=TRUE}
s_ins <- rownames(de_des_ins)
d_ins <- bulk_data_ins[, s_ins]

dds_ins <- DESeqDataSetFromMatrix(countData = d_ins,
                                  colData = de_des_ins,
                                  design = ~ tissue + insulin_stim)

de_des_ins$ind.n <- as.factor((c(1, 1, 1, 1, 2, 2, 2, 2)))

dds_ins <- DESeq(dds_ins)

resultsNames(dds_ins) # lists the coefficients


res_ins <- results(dds_ins, name="insulin_stim_1_vs_0")
res_ins_df <- as.data.frame(res_ins)
res_ins

resLFC_ins <- lfcShrink(dds_ins, coef="insulin_stim_1_vs_0", type="apeglm")
resLFC_ins
```


#### Plot results

```{r deseq_ins_plot1, fig.width=6, fig.height=6, fig.show='hold'}
#' MA-plot
plotMA(res_ins, ylim=c(-2,2), main = "MA-plot")
plotMA(resLFC_ins, main = "MA-plot with shrunken log2 fc (apeglm)")
```

```{r deseq_ins_restable_plot, fig.width=3, fig.height=2, fig.show='hold'}
res_ins_df <- as.data.frame(res_ins)

res_ins_head_up <- res_ins_df %>% 
  dplyr::arrange(padj) %>%
  dplyr::filter(padj<0.05 & log2FoldChange > 0) %>%
  head(5)
res_ins_head_down <- res_ins_df %>% 
  dplyr::arrange(padj) %>%
  dplyr::filter(padj<0.05 & log2FoldChange < 0) %>%
  head(5)
res_ins_head <- rbind(res_ins_head_up, res_ins_head_down)

d_plot <- t(d_ins)
d_plot <- cbind(d_plot, de_des_ins)

for(g in rownames(res_ins_head)){
  p <- ggplot(d_plot, aes_string(x = "insulin_stim", y = g)) +
    geom_boxplot() +
    geom_point(aes(color=subject_alias)) +
    scale_color_manual(values = colors_multi) +
    theme_classic()
  print(p)
}
```



```{r deseq_ins_vulcano, fig.width=6, fig.height=5, fig.show='hold'}
#' vulcano plot
res_ins_df$significant_0.05 <- res_ins_df$padj < 0.05 
res_ins_df$gene_labels <- ifelse(res_ins_df$padj < 0.05 & abs(res_ins_df$log2FoldChange) > 2, # & abs(res_ins_df$log2FoldChange) > 1, 
                                  rownames(res_ins_df), "")

glabs_ins <- res_ins_df[res_ins_df$gene_labels != "" & !is.na(res_ins_df$gene_labels), "gene_labels"]

minmax_xilm <- round( max(abs(res_ins_df$log2FoldChange)) + 1 )

p <- ggplot(res_ins_df, aes(x = log2FoldChange, y = -log10(padj), color = significant_0.05)) +
        # geom_hline(yintercept = -log10(0.05), linetype="dashed", color = colors_multi[3], alpha=.8) +
        geom_hline(yintercept = 0, linetype="solid", color = "grey90") +
        geom_vline(xintercept = c(-2, 2), linetype="dashed", color = "grey90") +
        geom_point(size = .5, alpha = 0.8) +
        geom_text_repel(data = res_ins_df, aes(x = log2FoldChange, y = -log10(padj), label = gene_labels), color = colors_multi[3], size = 3) +
        scale_color_manual(values = colors_multi[c(1,3)]) +
        xlim(c(-minmax_xilm, minmax_xilm)) +
        labs(title = "DE genes: Before vs after insulin") +
        theme_classic() +
        theme(legend.position = "none",
              plot.title = element_text(size = rel(1.5), hjust = 0.5),
              axis.title = element_text(size = rel(1.25)),
              axis.line = element_line(colour = "grey50"));p

fname <- "bulk_dea_insulin_vulcano"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 4, height = 4, useDingbats = F);p;dev.off()
tiff(filename = file.path(DIR_FIG, paste0(fname, ".tiff")), units = "cm", width = 4*2.5, height = 4*2.5, res = 600); p; dev.off()
```



```{r deseq_ins_plot2, fig.width=12, fig.height=6, fig.show='hold'}
#' vulcano plot
res_ins_df <- as.data.frame(res_ins_df)

res_ins_df$significant_0.05 <- res_ins_df$pvalue < 0.05 
res_ins_df$gene_labels <- ifelse(res_ins_df$significant_0.05 == T & abs(res_ins_df$log2FoldChange) > 1.5, 
                                 rownames(res_ins_df), "")

glabs_ins <- res_ins_df[res_ins_df$gene_labels != "" & !is.na(res_ins_df$gene_labels), "gene_labels"]

minmax_xilm <- round( max(abs(res_ins_df$log2FoldChange)) + 1 )

ggplot(res_ins_df, aes(x = log2FoldChange, y = -log10(pvalue), color = significant_0.05)) +
  geom_hline(yintercept = -log10(0.05), linetype="dashed", color = colors_main[2], alpha=.8) +
  geom_vline(xintercept = c(-2, 2), linetype="dashed", color = "grey90") +
  geom_point(size = .5, alpha = 0.8) +
  geom_text_repel(data = res_ins_df, aes(x = log2FoldChange, y = -log10(pvalue), label = gene_labels), color = colors_main[2], size = 3) +
  scale_color_manual(values = colors_main) +
  xlim(c(-minmax_xilm, minmax_xilm)) +
  labs(title = "DE genes: lean vs obese after insulin treatment") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        axis.title = element_text(size = rel(1.25)),
        axis.line = element_line(colour = NA))

```
  
*No adjusted p-vals was possible to calculate with this contrast, thus the unadjusted p-values are displayed here. Positive FC: higher expr in obese as a response to insulin. The horizontal yellow dashed line corresponds to -log10(0.05), and the vertical grey dashed lines correspond to log2FC = 2.*  
  
  
```{r deseq_ins_plot2_table}
#' table with exp values for genes labelled in above vulcano plot
de_des_ins[, c("tissue", "condition", "insulin_stim")]
datatable(d_ins[glabs_ins, rownames(de_des_ins)], rownames = T, options = list(scrollX = TRUE), caption = "Expression of labelled genes in above vulcano plot.")
```

<br><br>


#### Results table
  
View and export results table
  
```{r deseq_ins_res_table}
#' Table with exp values for genes labelled in above vulcano plot
head(res_ins_df[order(res_ins_df$padj), ])
```
  
  
```{r deseq_ins_res_export}
table_export <- res_ins_df[order(res_ins_df$padj), ]
table_export$gene <- rownames(table_export)
table_export <- table_export[, c("gene", colnames(table_export)[-7])]

fname <- "bulk_dea_insulin.tsv"
write.table(x = table_export, file = file.path(DIR_RES, "tables", fname), quote = F, sep = "\t")
# res_ins_df <- read.table(file = file.path(DIR_RES, "tables", "bulk_dea_insulin.tsv"), sep = "\t")
```



# Compare CAGE and insulin

## Vulcano plots

```{r cage_vulcano, fig.width=10, fig.height=4.5, fig.show='hold'}
#' CAGE
cage$significant_0.05 <- cage$FDR < 0.05 
cage$gene_labels <- ifelse(cage$FDR < 0.05 & abs(cage$logFC) > 1.8 & cage$symbol != "None",
                           rownames(cage),
                           "")

glabs_cage <- cage[cage$gene_labels != "" & !is.na(cage$gene_labels), "gene_labels"]
cage[glabs_cage, ]$gene_labels <- as.character(cage[glabs_cage, "symbol"])
minmax_xilm_cage <- 7


# Visium
res_ins_df$significant_0.05 <- res_ins_df$padj < 0.05 
res_ins_df$gene_labels <- ifelse(res_ins_df$padj < 0.05 & abs(res_ins_df$log2FoldChange) > 1.8, # & abs(res_ins_df$log2FoldChange) > 1, 
                                  rownames(res_ins_df), "")
glabs_ins <- res_ins_df[res_ins_df$gene_labels != "" & !is.na(res_ins_df$gene_labels), "gene_labels"]
minmax_xilm_visium <- round( max(abs(res_ins_df$log2FoldChange)) + 0.5 )


#' Plot
p1 <- ggplot(cage, aes(x = logFC*-1, y = -log10(FDR), color = significant_0.05)) +
          geom_hline(yintercept = 0, linetype="solid", color = "grey90") +
          geom_vline(xintercept = c(-2, 2), linetype="dashed", color = "grey90") +
          geom_point(size = .5, alpha = 0.8) +
          geom_text_repel(data = cage, aes(x = logFC*-1, y = -log10(FDR), label = gene_labels), color = colors_multi[3], size = 3) +
          scale_color_manual(values = colors_multi[c(1,3)]) +
          xlim(c(-minmax_xilm_cage, minmax_xilm_cage)) +
          labs(title = "CAGE") +
          theme_classic() +
          theme(legend.position = "none",
                plot.title = element_text(size = rel(1.5), hjust = 0.5),
                axis.title = element_text(size = rel(1.25)),
                axis.line = element_line(colour = "grey50"))

p2 <- ggplot(res_ins_df, aes(x = log2FoldChange, y = -log10(padj), color = significant_0.05)) +
        geom_hline(yintercept = 0, linetype="solid", color = "grey90") +
        geom_vline(xintercept = c(-2, 2), linetype="dashed", color = "grey90") +
        geom_point(size = .5, alpha = 0.8) +
        geom_text_repel(data = res_ins_df, aes(x = log2FoldChange, y = -log10(padj), label = gene_labels), color = colors_multi[3], size = 3) +
        scale_color_manual(values = colors_multi[c(1,3)]) +
        xlim(c(-minmax_xilm_visium, minmax_xilm_visium)) +
        labs(title = "Visium") +
        theme_classic() +
        theme(legend.position = "none",
              plot.title = element_text(size = rel(1.5), hjust = 0.5),
              axis.title = element_text(size = rel(1.25)),
              axis.line = element_line(colour = "grey50"))


p <- (p1+p2);p

fname <- "bulk_dea_insulin_cage_visium"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 10, height = 4.5, useDingbats = F);p;dev.off()
tiff(filename = file.path(DIR_FIG, paste0(fname, ".tiff")), units = "cm", width = 10*2.5, height = 4.5*2.5, res = 600); p; dev.off()
```
  
  
  
  
## DE list overlap

```{r cage_mod}
cage_mod <- cage_sign
cage_mod$logFC <- cage_mod$logFC*-1
```

```{r dea_insulin_up}
cage_up <- subset(cage_mod, logFC > 0 & FDR < 0.05); dim(cage_up)
visium_up <- subset(res_ins_df, log2FoldChange > 0 & padj < 0.05); dim(visium_up)
ins_up_visium <- rownames(visium_up)

ins_up_overlap1 <- intersect(rownames(visium_up), cage_up$symbol)
ins_up_overlap2 <- intersect(rownames(visium_up), cage_up$hgnc_symbol)
ins_up_overlap <- unique(c(ins_up_overlap1, ins_up_overlap2))
length(ins_up_overlap)
ins_up_overlap
```


```{r dea_insulin_down}
cage_down <- subset(cage_mod, logFC < 0 & FDR < 0.05); dim(cage_down)
visium_down <- subset(res_ins_df, log2FoldChange < 0 & padj < 0.05); dim(visium_down)
ins_down_visium <- rownames(visium_down)

ins_down_overlap1 <- intersect(rownames(visium_down), as.character(cage_down$symbol))
ins_down_overlap2 <- intersect(rownames(visium_down), as.character(cage_down$hgnc_symbol))
ins_down_overlap <- unique(c(ins_down_overlap1, ins_down_overlap2))
length(ins_down_overlap)
ins_down_overlap
```


```{r dea_insulin_down_view}
cage_mod$symbol_updated <- cage_mod$symbol
cage_mod[cage_mod$hgnc_symbol %in% c(ins_up_overlap2, ins_down_overlap2), "symbol_updated"] <- cage_mod[cage_mod$hgnc_symbol %in% c(ins_up_overlap2, ins_down_overlap2), "hgnc_symbol"]
```


<br><br>
  
  
### Joint table  
  
```{r dea_insulin_updown}
#' cage
cols_keep <- c("Annotation","logFC","logCPM","PValue","FDR","symbol","symbol_updated")
df_up_cage <- cage_mod[cage_mod$symbol_updated %in% ins_up_overlap, cols_keep]
df_up_cage$reg <- "up"
df_down_cage <- cage_mod[cage_mod$symbol_updated %in% ins_down_overlap, cols_keep]
df_down_cage$reg <- "down"
df_cage <- rbind(df_up_cage, df_down_cage) %>%
  dplyr::group_by(symbol_updated) %>%
  dplyr::filter(dplyr::row_number()==1)

df_cage$gene <- df_cage$symbol_updated
colnames(df_cage)[1:5] <- paste("cage", names(df_cage[1:5]), sep = ".")
df_cage <- as.data.frame(df_cage)
rownames(df_cage) <- df_cage$gene

#' visium
df_visium <- subset(res_ins_df[c(ins_up_overlap, ins_down_overlap), ])[, 2:6]
df_visium$reg <- "up"
df_visium[df_visium$log2FoldChange<0, "reg"] <- "down"
colnames(df_visium)[1:5] <- paste("visium", names(df_visium[1:5]), sep = ".")
df_visium

#' Join
df_cage_visium <-  merge(df_visium, df_cage[, c(grep("cage.", colnames(df_cage), value = T), "gene")], by = 'row.names')
df_cage_visium <- df_cage_visium[, c("gene", 
                                     "reg", 
                                     c(grep("visium.", colnames(df_visium), value = T)), 
                                     c(grep("cage.", colnames(df_cage), value = T)))]
rownames(df_cage_visium) <- df_cage_visium$gene
# df_cage_visium
```

```{r export_table_cagevisium}
fname <- "bulk_dea_ins_cage_visium.tsv"
write.table(x = df_cage_visium, file = file.path(DIR_RES, "tables", fname), quote = F, sep = "\t", row.names = F)
```


### Plot visium box

```{r plot_visium_ins_up}
d_up <- t(d_ins[ins_up_overlap, ])
d_up <- cbind(d_up, de_des_ins[,c("insulin_stim", "condition", "tissue")])
d_up$sample <- rownames(d_up)
names(d_up)[names(d_up) == 'insulin_stim'] <- 'Insulin'

d_up_long <- gather(d_up, gene, expression, ins_up_overlap[1]:ins_up_overlap[length(ins_up_overlap)], factor_key=TRUE)

p <- ggplot(d_up_long, aes(x = reorder(gene, expression, FUN = median), y = expression, fill = Insulin)) +
  geom_boxplot() +
  scale_fill_manual(values = colors_insulin) +
  # geom_point(size = 0.5, alpha = 0.7, aes(color=tissue)) +
  # scale_color_manual(values = colors_multi[5:9]) +
  scale_y_log10() +
  labs(x="", y="Summed expression", title = "Upregulated genes after insulin") +
  # rotate() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        plot.title = element_text(hjust = 0.5),
        text = element_text(size = 15), 
        panel.grid.major.x = element_line(colour="grey90"));p

fname <- "bulk_dea_insulin_cage_visium_up"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 10, height = 4.5, useDingbats = F);p;dev.off()
tiff(filename = file.path(DIR_FIG, paste0(fname, ".tiff")), units = "cm", width = 10*2.5, height = 4.5*2.5, res = 600); p; dev.off()
```


```{r plot_visium_ins_down}
d_down <- t(d_ins[ins_down_overlap, ])
d_down <- cbind(d_down, de_des_ins[,c("insulin_stim", "condition", "tissue")])
d_down$sample <- rownames(d_down)
names(d_down)[names(d_down) == 'insulin_stim'] <- 'Insulin'

d_down_long <- gather(d_down, gene, expression, ins_down_overlap[1]:ins_down_overlap[length(ins_down_overlap)], factor_key=TRUE)

p <- ggplot(d_down_long, aes(x = reorder(gene, expression, FUN = median), y = expression, fill = Insulin)) +
  geom_boxplot() +
  scale_fill_manual(values = colors_insulin) +
  # geom_point(size = 0.5, alpha = 0.7, aes(color=tissue)) +
  scale_color_manual(values = colors_multi[5:9]) +
  scale_y_log10() +
  labs(x="", y="Summed expression", title = "Downregulated genes after insulin") +
  # rotate() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        plot.title = element_text(hjust = 0.5),
        text = element_text(size = 15), 
        panel.grid.major.x = element_line(colour="grey90"));p

fname <- "bulk_dea_insulin_cage_visium_down"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 6, height = 4.5, useDingbats = F);p;dev.off()
tiff(filename = file.path(DIR_FIG, paste0(fname, ".tiff")), units = "cm", width = 6*2.5, height = 4.5*2.5, res = 600); p; dev.off()
```


### Plot Visium and CAGE

```{r plot_cage_ins_up}
p1 <- ggplot(df_cage_visium, aes(x=reorder(gene, visium.log2FoldChange), y=cage.logFC, fill=cage.FDR)) +
  geom_bar(stat="identity") +
  scale_fill_gradient(low = "grey30", high = "grey70") +
  labs(x="", y="log2 fold change", title = "CAGE") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        plot.title = element_text(hjust = 0.5));

p2 <- ggplot(df_cage_visium, aes(reorder(gene, visium.log2FoldChange), visium.log2FoldChange, fill=visium.padj)) +
  geom_bar(stat="identity") +
  scale_fill_gradient(low = "grey30", high = "grey70") +
  labs(x="", y="log2 fold change", title = "Visium") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        plot.title = element_text(hjust = 0.5));

p <- plot_grid(p1, p2, nrow = 2);p

fname <- "bulk_dea_insulin_cage_visium_log2fc"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 8, height = 8, useDingbats = F);p;dev.off()
tiff(filename = file.path(DIR_FIG, paste0(fname, ".tiff")), units = "cm", width = 8*2.5, height = 8*2.5, res = 600); p; dev.off()
```

  
  
# Cluster response of insulin-responding genes  
  
  
## Visium adipocyte data

### Gene expression  

```{r prep_visium_adipocyte_assaydata}
se_ins_data_all <- GetAssayData(subset(se_ins, cluster_anno %in% c("Adipocyte 1", "Adipocyte 2")), slot = "data")  # normalized data
```


#### All genes

```{r prep_visium_adipocyte_insulin_all}
se_ins_data_grouped_all <- data.frame(row.names = rownames(se_ins_data_all))
s_names <- unique(se_ins@meta.data$sample_name)  # metadata_ins$sample_name
sc_names <- c(paste(3, s_names, sep="."), paste(4, s_names, sep="."))
se_ins_metadata_all <- data.frame(row.names = sc_names,
                                  insulin_stim = rep(0, length(sc_names)), 
                                  bmi = rep(0, length(sc_names)), 
                                  condition = rep("", length(sc_names)), 
                                  subject_id = rep("", length(sc_names)), 
                                  m_value = rep(0, length(sc_names)), 
                                  cluster = rep("", length(sc_names)), 
                                  stringsAsFactors = F)

for(s in s_names){
  s_spots <- colnames(subset(se_ins, sample_name == s))
  for(c in c(4, 5)){
    c_spots <- colnames(subset(se_ins, seurat_clusters == c))
    sc_spots <- intersect(c_spots, s_spots)
    sc_data <- data.frame(Matrix::rowMeans(se_ins_data_all[, sc_spots]))
    sc_name <- paste(c, s, sep=".")
    colnames(sc_data) <- sc_name
    se_ins_data_grouped_all <- cbind(se_ins_data_grouped_all, sc_data)
    se_ins_metadata_all[sc_name, ] <- metadata_ins[s, c("insulin_stim", "bmi", "condition", "subject_id", "m_value")]
    if(c==5){
      se_ins_metadata_all[sc_name, "cluster"] <- "Adipocyte 1"  # LEP
    } else if (c==4){
      se_ins_metadata_all[sc_name, "cluster"] <- "Adipocyte 2"  # ADIPOQ
    }
  }
}

se_ins_metadata_all <- se_ins_metadata_all[colnames(se_ins_data_grouped_all), ]
se_ins_metadata_all
head(se_ins_data_grouped_all)
```



#### Insulin responding genes  

```{r prep_visium_adipocyte_insulin_selected}
genes_ins <- as.character(df_cage_visium$gene)
genes_ins_include <- intersect(rownames(se_ins_data_all), genes_ins)
se_ins_data <- as.data.frame(se_ins_data_all[genes_ins_include, ])

se_ins_data_grouped <- data.frame(row.names = rownames(se_ins_data))
s_names <- unique(se_ins@meta.data$sample_name)
sc_names <- c(paste(3, s_names, sep="."), paste(4, s_names, sep="."))
se_ins_metadata <- data.frame(row.names = sc_names,
                              insulin_stim = rep(0, length(sc_names)), 
                              bmi = rep(0, length(sc_names)), 
                              condition = rep("", length(sc_names)), 
                              subject_id = rep("", length(sc_names)), 
                              m_value = rep(0, length(sc_names)), 
                              cluster = rep("", length(sc_names)), 
                              stringsAsFactors = F)

for(s in s_names){
  s_spots <- colnames(subset(se_ins, sample_name == s))
  for(c in c(4, 5)){
    c_spots <- colnames(subset(se_ins, seurat_clusters == c))
    sc_spots <- intersect(c_spots, s_spots)
    sc_data <- data.frame(Matrix::rowMeans(se_ins_data[, sc_spots]))
    sc_name <- paste(c, s, sep=".")
    colnames(sc_data) <- sc_name
    se_ins_data_grouped <- cbind(se_ins_data_grouped, sc_data)
    se_ins_metadata[sc_name, ] <- metadata_ins[s, c("insulin_stim", "bmi", "condition", "subject_id", "m_value")]
    if(c==5){
      se_ins_metadata[sc_name, "cluster"] <- "Adipocyte 1"  # LEP
    } else if (c==4){
      se_ins_metadata[sc_name, "cluster"] <- "Adipocyte 2"  # ADIPOQ
    }
  }
}

# se_ins_data_grouped
se_ins_metadata <- se_ins_metadata[colnames(se_ins_data_grouped), ]
se_ins_metadata
head(se_ins_data_grouped)
```


#### Save tables

```{r save_grouped_ins_tables}
fname1 <- "bulk_grouped_cluster_expression-all_genes.tsv"
write.table(x = se_ins_data_grouped_all, file = file.path(DIR_RES, "tables", fname1), quote = F, sep = "\t", row.names = T)

fname2 <- "bulk_grouped_cluster_expression-insulin_resp_genes.tsv"
write.table(x = se_ins_data_grouped, file = file.path(DIR_RES, "tables", fname2), quote = F, sep = "\t", row.names = T)
```
<br><br>


### Group insulin gene response

```{r plot_visium_adipocyte_insulin_prep_score}
se_ins_data_score <- rbind(colMeans(se_ins_data_grouped[ins_up_overlap, ]), 
                           colMeans(se_ins_data_grouped[ins_down_overlap, ]))
rownames(se_ins_data_score) <- c("up", "down")
# se_ins_data_score


# Calculate difference score for each indvidual
se_ins_data_score_mean <- data.frame()
se_ins_data_score_ratio <- data.frame()
for(s in unique(se_ins_metadata$subject_id)){
  bool_plin_0 <- rownames(se_ins_metadata[(se_ins_metadata$subject_id==s & se_ins_metadata$cluster=="Adipocyte 2" & se_ins_metadata$insulin_stim=="0"),])
  bool_plin_1 <- rownames(se_ins_metadata[(se_ins_metadata$subject_id==s & se_ins_metadata$cluster=="Adipocyte 2" & se_ins_metadata$insulin_stim=="1"),])
  bool_lep_0 <- rownames(se_ins_metadata[(se_ins_metadata$subject_id==s & se_ins_metadata$cluster=="Adipocyte 1" & se_ins_metadata$insulin_stim=="0"),])
  bool_lep_1 <- rownames(se_ins_metadata[(se_ins_metadata$subject_id==s & se_ins_metadata$cluster=="Adipocyte 1" & se_ins_metadata$insulin_stim=="1"),])
  
  pc <- .01 # pseudo count
  if(s=="NK49"){
    s_plin_all <- data.frame("ratio" = c(rowMeans(se_ins_data_grouped[ins_up_overlap, bool_plin_1]+pc) / (se_ins_data_grouped[ins_up_overlap, bool_plin_0]+pc), 
                                           rowMeans(se_ins_data_grouped[ins_down_overlap, bool_plin_1]+pc) / (se_ins_data_grouped[ins_down_overlap, bool_plin_0]+pc)), 
                               row.names = c(ins_up_overlap, ins_down_overlap)
                               )

    s_lep_all <- data.frame("ratio" = c(rowMeans(se_ins_data_grouped[ins_up_overlap, bool_lep_1]+pc) / (se_ins_data_grouped[ins_up_overlap, bool_lep_0]+pc), 
                                       rowMeans(se_ins_data_grouped[ins_down_overlap, bool_lep_1]+pc) / (se_ins_data_grouped[ins_down_overlap, bool_lep_0]+pc)), 
                           row.names = c(ins_up_overlap, ins_down_overlap)
                           )

  } else {
    s_plin_all <- data.frame("ratio" = c((se_ins_data_grouped[ins_up_overlap, bool_plin_1]+pc) / (se_ins_data_grouped[ins_up_overlap, bool_plin_0]+pc), 
                                           (se_ins_data_grouped[ins_down_overlap, bool_plin_1]+pc) / (se_ins_data_grouped[ins_down_overlap, bool_plin_0]+pc)), 
                           row.names = c(ins_up_overlap, ins_down_overlap)
                           )
    s_lep_all <- data.frame("ratio" = c((se_ins_data_grouped[ins_up_overlap, bool_lep_1]+pc) / (se_ins_data_grouped[ins_up_overlap, bool_lep_0]+pc), 
                                        (se_ins_data_grouped[ins_down_overlap, bool_lep_1]+pc) / (se_ins_data_grouped[ins_down_overlap, bool_lep_0]+pc)), 
                       row.names = c(ins_up_overlap, ins_down_overlap)
                       )
  }
  s_plin <- data.frame("up" = mean(s_plin_all[ins_up_overlap, "ratio"]), 
                       "down" = mean(s_plin_all[ins_down_overlap, "ratio"]))
  s_lep <- data.frame("up" = mean(s_lep_all[ins_up_overlap, "ratio"]), 
                       "down" = mean(s_lep_all[ins_down_overlap, "ratio"]))
  
  rownames(s_plin) <- colnames(s_plin_all) <- paste0("PLIN.",s)
  rownames(s_lep) <- colnames(s_lep_all) <- paste0("LEP.",s)
  s_out <- rbind(s_plin, s_lep)
  
  if(dim(se_ins_data_score_ratio)[1]==0){
    se_ins_data_score_ratio <- cbind(s_plin_all, s_lep_all)
  } else {
    se_ins_data_score_ratio <- cbind(se_ins_data_score_ratio, cbind(s_plin_all, s_lep_all))
  }
  se_ins_data_score_mean <- rbind(se_ins_data_score_mean, s_out)
}

sd_up <- colSds(as.matrix(log2(se_ins_data_score_ratio[ins_up_overlap, ])))
sd_down <- colSds(as.matrix(log2(se_ins_data_score_ratio[ins_down_overlap, ])))
se_ins_data_score_log2sd <- data.frame("up" = sd_up,
                                       "down" = sd_down, row.names = colnames(se_ins_data_score_ratio))
```



#### Plot

```{r plot_visium_adipocyte_insulin_hm_setup}
my_gene_col <- data.frame("regulation" = df_cage_visium[genes_ins_include, c("reg")], 
                          row.names = rownames(df_cage_visium[genes_ins_include,]))
my_sample_col <- se_ins_metadata[, c("insulin_stim", 
                                     "condition",
                                     "subject_id",
                                     "m_value",
                                     "cluster")]
my_sample_col$insulin_stim <- as.character(my_sample_col$insulin_stim)

m_vals <- as.character(sort(as.numeric(unique(my_sample_col$m_value))))
my_colour = list(
    condition = c(lean = colors_main[1], obese = colors_main[2]),
    insulin_stim = c(`0` = colors_insulin[1], `1` = colors_insulin[2]),
    subject_id = c("NK49" =  colors_multi[1],
                   "NK50" =  colors_multi[2],
                   "NO137" = colors_multi[3],
                   "NO148" = colors_multi[4]),
    m_value = colors_mval,
    cluster = colors_adi[1:2],
             # c("Adipocyte 2" = "#907198",
             #  "Adipocyte 1" = "#BFABC3"),
    regulation = c(down = color_low2, up = color_high)
)

```


```{r plot_visium_adipocyte_insulin_hm_score_metadata_setup}
my_sample_col_2 <- my_sample_col %>%
  group_by(subject_id) %>%
  dplyr::top_n(n=1, wt = insulin_stim)

my_sample_col_2 <- my_sample_col_2[,-1]
my_sample_col_2 <- as.data.frame(my_sample_col_2)

my_sample_col_2$rname <- paste0("LEP.", my_sample_col_2[, "subject_id"])
for(c in unique(my_sample_col_2$cluster)){
  rows_incl <- rownames(my_sample_col_2[my_sample_col_2$cluster==c, ]) 
  if(c=="Adipocyte 2") {
    my_sample_col_2[rows_incl, "rname"] <- paste0("PLIN.", my_sample_col_2[rows_incl, "subject_id"])
  }
}
# se_ins_data_score_mean
my_sample_col_2 <- my_sample_col_2[!duplicated(my_sample_col_2$rname), ]
rownames(my_sample_col_2) <- my_sample_col_2$rname
my_sample_col_2

```


#### Heatmap  

##### All genes  

```{r plot_visium_adipocyte_insulin_hm, fig.width=6, fig.height=8, fig.show='hold'}
# DoHeatmap(subset(se_ins, seurat_clusters %in% c(3, 4)), features = as.character(df_cage_visium$gene)) +
#   scale_fill_gradientn(colours = rev(RColorBrewer::brewer.pal(5, "RdYlBu")))
# VlnPlot(se_ins, features = c("ADIPOQ", "LEP"))

cal_z_score <- function(x){
  (x - mean(x)) / sd(x)
}
 
data_hm_norm <- t(apply(se_ins_data_grouped, 1, cal_z_score))

fname <- "bulk_dea_insulin_cage_visium-cluster_heatmap_zscore"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 6, height = 8, useDingbats = F);
breaksList <- seq(-1.5, 1.5, by = 0.1)
p_hm <- pheatmap(data_hm_norm,  # se_ins_data_grouped, data_hm_norm
                 annotation_colors = my_colour,
                 annotation_col = my_sample_col[,c(1,3,4,5)],
                 annotation_row = my_gene_col,
                 breaks = breaksList,
                 # color = viridis::viridis(length(breaksList)),
                 color = colorRampPalette(c(color_low2, "white", color_high))(length(breaksList)-1),
                 border_color = "white"
                 )
dev.off()

# tiff(filename = file.path(DIR_FIG, paste0(fname, ".tiff")), units = "cm", width = 10*2.5, height = 8*2.5, res = 600)
# dev.off()
```



##### Ratio

```{r plot_visium_adipocyte_insulin_hm_score_all_savetable}
fname <- "bulk_grouped_cluster_expression_log2ratios-insulin_genes.tsv"
t_export <- log2(se_ins_data_score_ratio)
t_export$gene <- rownames(t_export)
write.table(x = t_export, file = file.path(DIR_RES, "tables", fname), quote = F, sep = "\t", row.names = F)
```


```{r plot_visium_adipocyte_insulin_hm_score_all, fig.width=4.5, fig.height=8}
fname <- "bulk_dea_insulin_cage_visium-cluster_heatmap_log2ratio"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 4.6, height = 8, useDingbats = F);

ph_plot <- log2(se_ins_data_score_ratio)
ph_plot <- ph_plot[,rownames(my_sample_col_2[order(my_sample_col_2$cluster, my_sample_col_2$m_value),])]
data_hm_norm <- t(apply(se_ins_data_grouped, 1, cal_z_score))

breaksList <- seq(-4, 4, by = 0.5)
pheatmap(ph_plot,
         annotation_colors = my_colour,
         annotation_col = my_sample_col_2[order(my_sample_col_2$cluster, my_sample_col_2$m_value), c(3,4)],
         annotation_row = my_gene_col, 
         cluster_rows = T,
         cluster_cols = F,
         gaps_col = 4,
         breaks = breaksList,
         color = colorRampPalette(c(color_low2, "white", color_high))(length(breaksList)-1),
         border_color = "white"
         )

dev.off()
```



```{r plot_visium_adipocyte_insulin_hm_score_mean}
fname <- "bulk_dea_insulin_cage_visium-cluster_heatmap_log2ratio_mean"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 6.5, height = 3, useDingbats = F);
breaksList <- seq(-2.5, 2.5, by = 0.5)
pheatmap(t(log2(se_ins_data_score_mean)),  # se_ins_data_grouped, data_hm_norm
         annotation_colors = my_colour,
         annotation_col = my_sample_col_2[,c(3,4)],
         cluster_rows = F,
         breaks = breaksList,
         color = colorRampPalette(c(color_low2, "white", color_high))(length(breaksList)-1),
         border_color = "white"
         )
dev.off()
```



#### Boxplot  

```{r plot_visium_adipocyte_insulin_box_score_dplot}
d_plot <- cbind(log2(se_ins_data_score_mean), my_sample_col_2)

fname <- "bulk_grouped_cluster_expression_log2ratios-insulin_genes_grouped.tsv"
t_export <- d_plot
write.table(x = t_export, file = file.path(DIR_RES, "tables", fname), quote = F, sep = "\t", row.names = F)
```


```{r plot_visium_adipocyte_insulin_box_score, fig.width = 4.5, fig.height = 2.5}
p1 <- ggplot(d_plot, aes(x=cluster, y=up))+
  geom_boxplot() +
  # geom_line(aes(group=subject_id), color = "grey80") +
  geom_point(size=2, aes(color=as.factor(m_value))) +
  scale_color_manual(values = my_colour$m_value) +
  geom_hline(yintercept = 0, linetype="solid", color = "grey50") +
  ylim(min(d_plot[,1:2]), max(d_plot[,1:2])) +
  labs(y="Expression log2-ratio") +
  labs(title="Upreg. genes", x="") +
  theme_classic() +
  theme(legend.position = "none");p1
  
p2 <- ggplot(d_plot, aes(x=cluster, y=down)) +
  geom_boxplot() +
  # geom_line(aes(group=subject_id), color = "grey80") +
  geom_point(size=2, aes(color=as.factor(m_value))) +
  scale_color_manual(values = my_colour$m_value) +
  geom_hline(yintercept = 0, linetype="solid", color = "grey50") +
  # facet_wrap(~condition) +
  ylim(min(d_plot[,1:2]), max(d_plot[,1:2])) +
  labs(y="Expression log2-ratio") +
  labs(title="Downreg. genes", x="") +
  theme_classic() +
  theme(legend.position = "none");p2

fname <- "bulk_dea_insulin_cage_visium-cluster_box_scores"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 5, height = 2.5, useDingbats = F);
plot_grid(p1, p2, nrow=1)
dev.off()
```

<br>

#### Corr plot

```{r plot_visium_adipocyte_insulin_corr_score, fig.width = 4.5, fig.height = 2.5}
expr_vals <- c(d_plot$up, d_plot$down)
d_plot2 <- rbind(d_plot[,c(3:7)], d_plot[,c(3:7)])
d_plot2$log2_ratio <- expr_vals
d_plot2$direction <- "down"
d_plot2[1:8, "direction"] <- "up"
d_plot2$m_value <- as.character(d_plot2$m_value)
# d_plot2$sd <- c(se_ins_data_score_sd$up, se_ins_data_score_sd$down)

p1 <- ggplot(subset(d_plot2, cluster=="Adipocyte 1"), aes(x=as.numeric(m_value), y=log2_ratio)) +
  geom_line(aes(group=direction), color = "grey80") +
  geom_point(size=2, aes(color=as.factor(m_value))) +
  scale_color_manual(values = my_colour$m_value) +
  geom_hline(yintercept = 0, linetype="solid", color = "grey50") +
  ylim(min(d_plot2$log2_ratio), max(d_plot2$log2_ratio)) +
  labs(y="Expression log2-ratio") +
  labs(title="Adipocyte 1", x="m-value") +
  theme_classic() +
  theme(legend.position = "none")

p2 <- ggplot(subset(d_plot2, cluster=="Adipocyte 2"), aes(x=as.numeric(m_value), y=log2_ratio)) +
  geom_line(aes(group=direction), color = "grey80") +
  geom_point(size=2, aes(color=as.factor(m_value))) +
  scale_color_manual(values = my_colour$m_value) +
  geom_hline(yintercept = 0, linetype="solid", color = "grey50") +
  ylim(min(d_plot2$log2_ratio), max(d_plot2$log2_ratio)) +
  labs(y="Expression log2-ratio") +
  labs(title="Adipocyte 2", x="m-value") +
  theme_classic() +
  theme(legend.position = "none")

fname <- "bulk_dea_insulin_cage_visium-cluster_corr_scores"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 5, height = 2.5, useDingbats = F);
plot_grid(p1, p2, nrow=1)
dev.off()
```



#### T test

Paired t-test between clusters

```{r ttest_visium_adipocyte_insulin_score}
print("UP")
x_LEP <- d_plot[d_plot$cluster=="Adipocyte 1","up"]
y_PLIN <- d_plot[d_plot$cluster!="Adipocyte 1","up"]
t.test(x_LEP, y_PLIN, paired = TRUE, alternative = "two.sided")
t.test(x_LEP, y_PLIN, paired = TRUE, alternative = "less")


print("DOWN")
x_LEP <- d_plot[d_plot$cluster=="Adipocyte 1","down"]
y_PLIN <- d_plot[d_plot$cluster!="Adipocyte 1","down"]
t.test(x_LEP, y_PLIN, paired = TRUE, alternative = "two.sided") 
t.test(x_LEP, y_PLIN, paired = TRUE, alternative = "greater") 
```


#### Barplot  

```{r plot_visium_adipocyte_insulin_bars}
#' Create plot data
df_cage_visium_adi <- rbind(df_cage_visium[1:2], df_cage_visium[1:2])
df_cage_visium_adi$cluster <- "Adipocyte 1"
df_cage_visium_adi[1:dim(df_cage_visium)[1], "cluster"] <- "Adipocyte 2"

df_cage_visium_adi <- merge(df_cage_visium_adi, se_ins_data_score_ratio)
s_ids <- unique(metadata_ins$subject_id)
for(s in s_ids){
  col_select <- grep(pattern = s, x = colnames(se_ins_data_score_ratio), value = T)
  s_data <- se_ins_data_score_ratio[, col_select]
  s_data$gene <- rownames(s_data)
  
  s_new <- data.frame(subject = c(s_data[, col_select[1]], 
                                  s_data[, col_select[2]]),
                      cluster = c(rep(x = strsplit(col_select[1], split = "\\.")[[1]][1], dim(s_data)[1]),
                                  rep(x = strsplit(col_select[2], split = "\\.")[[1]][1], dim(s_data)[1])),
                      gene = c(s_data$gene, s_data$gene),
                      stringsAsFactors = F)
  colnames(s_new) <- c(s, "cluster", "gene")
  s_new[s_new$cluster == "LEP", "cluster"] <- "Adipocyte 1"
  s_new[s_new$cluster == "PLIN", "cluster"] <- "Adipocyte 2"
  df_cage_visium_adi <- merge(df_cage_visium_adi, s_new, by=c("gene", "cluster"))
}

df_cage_visium_adi[, s_ids] <- log2(df_cage_visium_adi[, s_ids])
df_cage_visium_adi$ratio_mean <- rowMeans(df_cage_visium_adi[, s_ids])
df_cage_visium_adi$ratio_sd <- rowSds(as.matrix(df_cage_visium_adi[, s_ids]))


#' Plot
colors_fill <- my_colour$cluster
# names(colors_fill) <- c("Adipocyte 2", "Adipocyte 1")

# All:
p1 <- ggplot(df_cage_visium_adi, aes(x=reorder(gene, ratio_mean), y=ratio_mean, fill=cluster)) +
  geom_bar(stat="identity", color="black", position=position_dodge(), width=0.72) +
  geom_hline(yintercept = 0, linetype="solid", color = "black") +
  geom_errorbar(aes(ymin=ratio_mean-ratio_sd, ymax=ratio_mean+ratio_sd), width=.25,
                 position=position_dodge(.8)) +
  labs(x="", y="Log2-ratio", title = "") +
  scale_fill_manual(values=colors_fill) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0.5));p1

fname <- "bulk_dea_insulin_cage_visium-cluster_ratios_bar"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 12, height = 4, useDingbats = F);p1;dev.off()
tiff(filename = file.path(DIR_FIG, paste0(fname, ".tiff")), units = "cm", width = 12*2.5, height = 4*2.5, res = 600); p1; dev.off()


# Only NO148
p2 <- ggplot(df_cage_visium_adi, aes(x=reorder(gene, NO148), y=NO148, fill=cluster)) +
  geom_bar(stat="identity", color="black", position=position_dodge(), width=0.72) +
  geom_hline(yintercept = 0, linetype="solid", color = "black") +
  labs(x="", y="Log2-ratio", title = "NO148") +
  scale_fill_manual(values=colors_fill) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0.5));p2

fname <- "bulk_dea_insulin_cage_visium-cluster_ratios_bar_NO148"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 12, height = 4, useDingbats = F);p2;dev.off()
```



<br><br>

## Visium: all cluster data

### Gene expression  
```{r data_visium_adipocyte_insulin_allclust}
se_ins_allc <- GetAssayData(se_ins[genes_ins_include,], slot = "data")  # normalized data for genes of interest
```
  
  
#### Insulin responding genes 

```{r prep_visium_adipocyte_insulin_allclust}
se_ins_allc_grouped <- data.frame(row.names = rownames(se_ins_allc))
s_names <- unique(se_ins@meta.data$sample_name)  # metadata_ins$sample_name
c_ids <- as.character(sort(as.numeric(as.character(unique(se_ins@meta.data$seurat_clusters)))))
sc_names <- c()
for(c in c_ids){
  sc_names <- c(sc_names, paste(c, s_names, sep=".") )
}

se_ins_metadata_allc <- data.frame(row.names = sc_names,
                                  insulin_stim = rep(0, length(sc_names)), 
                                  bmi = rep(0, length(sc_names)), 
                                  condition = rep("", length(sc_names)), 
                                  subject_id = rep("", length(sc_names)), 
                                  m_value = rep(0, length(sc_names)), 
                                  cluster = rep("", length(sc_names)), 
                                  stringsAsFactors = F)

for(s in s_names){
  for(c in c_ids){
      s_spots <- colnames(subset(se_ins, sample_name == s))
      c_spots <- colnames(subset(se_ins, seurat_clusters == c))
      sc_spots <- intersect(c_spots, s_spots)
      
      spotn_cutoff <- 10  # originally 1
      if (length(sc_spots) >= spotn_cutoff) {
        sc_data <- data.frame(Matrix::rowMeans(se_ins_allc[, sc_spots]))
      } else if (length(sc_spots) == 0) {
        # sc_data <- data.frame(se_ins_allc[, sc_spots])  # won't work if no spots
        # sc_data[,1] <- 0
        sc_data <- data.frame(value = rep(0, nrow(se_ins_allc)), row.names = rownames(se_ins_allc))
      } else if (length(sc_spots) == 1) {
        sc_data <- data.frame(se_ins_allc[, sc_spots])
        sc_data[,1] <- 0
      } else {
        sc_data <- data.frame(Matrix::rowMeans(se_ins_allc[, sc_spots]))
        sc_data[,1] <- 0
      }
      # print(paste(c, s, sep="."))
      # print(head(sc_data))
      
      sc_name <- paste(c, s, sep=".")
      colnames(sc_data) <- sc_name
      se_ins_allc_grouped <- cbind(se_ins_allc_grouped, sc_data)
      se_ins_metadata_allc[sc_name, ] <- metadata_ins[s, c("insulin_stim", "bmi", "condition", "subject_id", "m_value")]
      se_ins_metadata_allc[sc_name, "cluster"] <- paste0("C", c)
  }
}

se_ins_metadata_allc <- se_ins_metadata_allc[colnames(se_ins_allc_grouped), ]
se_ins_metadata_allc
head(se_ins_allc_grouped)
```


```{r prep_visium_adipocyte_insulin_allclust_score}
se_ins_data_score_allc <- rbind(colMeans(se_ins_allc_grouped[ins_up_overlap, ]), 
                           colMeans(se_ins_allc_grouped[ins_down_overlap, ]))
rownames(se_ins_data_score_allc) <- c("up", "down")
# se_ins_data_score_allc


# Calculate difference score for each indvidual
se_ins_data_score_allc_mean <- data.frame()
se_ins_data_score_allc_ratio <- data.frame()
for(s in unique(se_ins_metadata_allc$subject_id)){
  pc <- .01 # pseudo count
  for(c in c_ids){
    bool_c_0 <- rownames(se_ins_metadata_allc[(se_ins_metadata_allc$subject_id==s & 
                                                 se_ins_metadata_allc$cluster==paste0("C",c) & 
                                                 se_ins_metadata_allc$insulin_stim=="0"),])
    bool_c_1 <- rownames(se_ins_metadata_allc[(se_ins_metadata_allc$subject_id==s & 
                                                 se_ins_metadata_allc$cluster==paste0("C",c) & 
                                                 se_ins_metadata_allc$insulin_stim=="1"),])
    
    up_sum_1 <- sum(se_ins_allc_grouped[ins_up_overlap, bool_c_1])
    up_sum_0 <- sum(se_ins_allc_grouped[ins_up_overlap, bool_c_0])
    down_sum_1 <- sum(se_ins_allc_grouped[ins_down_overlap, bool_c_1])
    down_sum_0 <- sum(se_ins_allc_grouped[ins_down_overlap, bool_c_0])
    
    
    ngenes_up_1 <- sum(se_ins_allc_grouped[ins_up_overlap, bool_c_1]>0)
    ngenes_up_0 <- sum(se_ins_allc_grouped[ins_up_overlap, bool_c_0]>0)
    ngenes_down_1 <- sum(se_ins_allc_grouped[ins_down_overlap, bool_c_1]>0)
    ngenes_down_0 <- sum(se_ins_allc_grouped[ins_down_overlap, bool_c_0]>0)
    
    if(s=="NK49"){
    s_c_all <- data.frame("ratio" = c(rowMeans(se_ins_allc_grouped[ins_up_overlap, bool_c_1]+pc) / (se_ins_allc_grouped[ins_up_overlap, bool_c_0]+pc), 
                                      rowMeans(se_ins_allc_grouped[ins_down_overlap, bool_c_1]+pc) / (se_ins_allc_grouped[ins_down_overlap, bool_c_0]+pc)), 
                          row.names = c(ins_up_overlap, ins_down_overlap)
                          )
    } else {
      s_c_all <- data.frame("ratio" = c((se_ins_allc_grouped[ins_up_overlap, bool_c_1]+pc) / (se_ins_allc_grouped[ins_up_overlap, bool_c_0]+pc), 
                                        (se_ins_allc_grouped[ins_down_overlap, bool_c_1]+pc) / (se_ins_allc_grouped[ins_down_overlap, bool_c_0]+pc)), 
                            row.names = c(ins_up_overlap, ins_down_overlap)
                            )
    }
    
    if (up_sum_1==0 | up_sum_0==0) {
      s_c_all[ins_up_overlap, "ratio"] <- 1
    } 
    if (down_sum_1==0 | down_sum_0==0){
      s_c_all[ins_down_overlap, "ratio"] <- 1
    }
    
    gene_cutoff <- 2
    if (ngenes_up_1 < gene_cutoff | ngenes_up_0 < gene_cutoff) {
      s_c_all[ins_up_overlap, ] <- 1
    } 
    if (ngenes_down_1 < gene_cutoff | ngenes_down_0 < gene_cutoff){
      s_c_all[ins_down_overlap, ] <- 1
    }
    
    s_clust <- data.frame("up" = mean(s_c_all[ins_up_overlap, "ratio"]), 
                          "down" = mean(s_c_all[ins_down_overlap, "ratio"]))

    rownames(s_clust) <- colnames(s_c_all) <- paste0(c, ".", s)
    
    if(dim(se_ins_data_score_allc_ratio)[1]==0){
      se_ins_data_score_allc_ratio <- s_c_all
    } else {
      se_ins_data_score_allc_ratio <- cbind(se_ins_data_score_allc_ratio, s_c_all)
    }
    se_ins_data_score_allc_mean <- rbind(se_ins_data_score_allc_mean, s_clust)
  }
}
```





```{r prep_visium_adipocyte_insulin_allclust_score_savetable}
fname <- "bulk_grouped_cluster_expression-all_clusters_grouped_mean_expr.tsv"
t_export <- se_ins_allc_grouped
t_export$gene <- rownames(t_export)
write.table(x = t_export, file = file.path(DIR_RES, "tables", fname), quote = F, sep = "\t", row.names = F)


fname <- "bulk_grouped_cluster_expression_log2ratios-all_clusters_heatmap_log2ratio.tsv"
t_export <- log2(se_ins_data_score_allc_ratio)
t_export$gene <- rownames(t_export)
write.table(x = t_export, file = file.path(DIR_RES, "tables", fname), quote = F, sep = "\t", row.names = F)


fname <- "bulk_grouped_cluster_expression_log2ratios-all_clusters_heatmap_log2ratio_mean.tsv"
t_export <- log2(se_ins_data_score_allc_mean)
t_export$cluster_subject <- rownames(t_export)
write.table(x = t_export, file = file.path(DIR_RES, "tables", fname), quote = F, sep = "\t", row.names = F)
```
  
  
  

### Heatmap

```{r hm_all_clus_colors}
cols_allc <- rep("grey80", length(c_ids))
names(cols_allc) <- paste0("C", c_ids)
cols_allc[c("C5", "C4")] <- colors_adi[c("Adipocyte 2", "Adipocyte 1")] # c("#F768A1", "#FBB4B9")

my_colour_allc = list(
    condition = c(lean = colors_main[1], obese = colors_main[2]),
    insulin_stim = c(`0` = colors_insulin[1], `1` = colors_insulin[2]),
    subject_id = c("NK49" =  colors_multi[3],
                   "NK50" =  colors_multi[4],
                   "NO137" = colors_multi[5],
                   "NO148" = colors_multi[6]),
    m_value = colors_mval,
    cluster = cols_allc,
    regulation = c(down = color_low, up = color_high)
)
```


```{r plot_visium_adipocyte_insulin_hm_score_metadata_allclust}
my_gene_col <- data.frame("regulation" = df_cage_visium[genes_ins_include, c("reg")], 
                          row.names = rownames(df_cage_visium[genes_ins_include,]))

my_sample_col_allc <- se_ins_metadata_allc[, c("insulin_stim", 
                                               "condition",
                                               "subject_id",
                                               "m_value",
                                               "cluster")]
my_sample_col_allc$insulin_stim <- as.character(my_sample_col_allc$insulin_stim)

my_sample_col_allc <- my_sample_col_allc %>%
  group_by(subject_id) %>%
  dplyr::top_n(n=1, wt = insulin_stim)

my_sample_col_allc <- my_sample_col_allc[,-1]
my_sample_col_allc <- as.data.frame(my_sample_col_allc)

my_sample_col_allc$rname <- paste0("0.", my_sample_col_allc[, "subject_id"])
for(c in c_ids){
  rows_incl <- rownames(my_sample_col_allc[my_sample_col_allc$cluster==paste0("C",c), ]) 
  my_sample_col_allc[rows_incl, "rname"] <- paste0(c, ".", my_sample_col_allc[rows_incl, "subject_id"])
}

# se_ins_data_score_mean
my_sample_col_allc <- my_sample_col_allc[!duplicated(my_sample_col_allc$rname), ]
rownames(my_sample_col_allc) <- my_sample_col_allc$rname
my_sample_col_allc
```


```{r plot_hm_all_clusters_log2ratio, fig.width=16, fig.height=8}
fname <- "bulk_dea_insulin_cage_visium-all_clusters_heatmap_log2ratio"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 16, height = 8, useDingbats = F);
breaksList <- seq(-2, 2, by = 0.01)
pheatmap(
   mat = (log2(se_ins_data_score_allc_ratio)),  # se_ins_data_grouped, data_hm_norm
   annotation_colors = my_colour_allc,
   annotation_col = my_sample_col_allc[,c(2,3,4)],
   annotation_row = my_gene_col,
   cluster_rows = T,
   breaks = breaksList,
   color = colorRampPalette(c(color_low2, "white", color_high))(length(breaksList)-1),
   border_color = "white"
   )
 dev.off()
```


```{r plot_visium_adipocyte_insulin_hm_score_mean_allc, fig.width=4, fig.height=12}
fname <- "bulk_dea_insulin_cage_visium-all_clusters_heatmap_log2ratio_mean"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 4.5, height = 12, useDingbats = F)
breaksList <- seq(-2.5, 2.5, by = 0.1)
pheatmap((log2(se_ins_data_score_allc_mean)),  # se_ins_data_grouped, data_hm_norm
         annotation_colors = my_colour_allc,
         annotation_row = my_sample_col_allc[,c(2,3,4)],
         # annotation_row = my_gene_col,
         cluster_rows = T,
         cluster_cols = F,
         breaks = breaksList,
         color = colorRampPalette(c(color_low2, "white", color_high))(length(breaksList)-1),
         border_color = "white"
         )
dev.off()
```


Only adipocytes PLIN and LEP  

```{r plot_visium_adipocyte_insulin_hm_score_mean, fig.width=4.5, fig.height=8}

ph_plot <- log2(se_ins_data_score_allc_ratio)
ph_plot <- ph_plot[, c(grep("^5.", colnames(ph_plot), value = T), grep("^4.", colnames(ph_plot), value = T))]
colnames(ph_plot) <- sub("^5.", "LEP.", colnames(ph_plot))
colnames(ph_plot) <- sub("^4.", "PLIN.", colnames(ph_plot))

ph_plot <- ph_plot[, rownames(my_sample_col_2[order(my_sample_col_2$cluster, my_sample_col_2$m_value),]) ]


fname <- "bulk_dea_insulin_cage_visium-all_cluster_heatmap_log2ratio_adipocytes"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 4.6, height = 8, useDingbats = F);

breaksList <- seq(-4, 4, by = 0.5)
pheatmap(ph_plot,
         annotation_colors = my_colour,
         annotation_col = my_sample_col_2[order(my_sample_col_2$cluster, my_sample_col_2$m_value), c(3,4)],
         annotation_row = my_gene_col, 
         cluster_rows = T,
         cluster_cols = F,
         gaps_col = 4,
         breaks = breaksList,
         color = colorRampPalette(c(color_low2, "white", color_high))(length(breaksList)-1),
         border_color = "white"
         )
dev.off()
```



### Box plot

```{r plot_visium_adipocyte_insulin_box_score_allc, fig.width=12, fig.height=4}
d_plot <- cbind(log2(se_ins_data_score_allc_mean), my_sample_col_allc)
d_plot$order <- rep(seq(1:19), 4)
d_plot[d_plot$up==0,"up"] <- NA
d_plot[d_plot$down==0,"down"] <- NA

p1 <- ggplot(d_plot, aes(x=reorder(cluster, order) , y=up))+
  geom_hline(yintercept = c(seq(-3,3,1),4), linetype="solid", color = "grey90") +
  geom_hline(yintercept = 0, linetype="solid", color = "grey50") +
  geom_boxplot() +
  geom_line(aes(group=subject_id), color = "grey80", alpha=0.6) +
  geom_point(size=2, aes(color=as.factor(m_value)), alpha=0.8) +
  geom_hline(yintercept = mean(d_plot$up, na.rm=T), linetype="dashed", color = "orange", alpha=0.8) +
  scale_color_manual(values = my_colour$m_value) +
  ylim(min(d_plot[,1:2])-.02, max(d_plot[,1:2])+0.02) +
  labs(y="Expression log2-ratio") +
  labs(title="Upreg. genes", x="Cluster", color="M value") +
  theme_classic() +
  theme(legend.position = "bottom")

p2 <- ggplot(d_plot, aes(x=reorder(cluster, order), y=down)) +
  geom_hline(yintercept = c(seq(-3,3,1),4), linetype="solid", color = "grey90") +
  geom_hline(yintercept = 0, linetype="solid", color = "grey50") +
  geom_boxplot() +
  geom_line(aes(group=subject_id), color = "grey80", alpha=0.6) +
  geom_point(size=2, aes(color=as.factor(m_value)), alpha=0.8) +
  geom_hline(yintercept = mean(d_plot$down, na.rm=T), linetype="dashed", color = "orange", alpha=0.8) +
  scale_color_manual(values = my_colour$m_value) +
  ylim(min(d_plot[,1:2])-.02, max(d_plot[,1:2])+0.02) +
  labs(y="Expression log2-ratio") +
  labs(title="Downreg. genes", x="Cluster", color="M value") +
  theme_classic() +
  theme(legend.position = "bottom")

fname <- "bulk_dea_insulin_cage_visium-all_cluster_box_scores"
pdf(file = file.path(DIR_FIG, paste0(fname, ".pdf")), width = 12, height = 4, useDingbats = F);
plot_grid(p1, p2, nrow=1)
dev.off()
```


```{r plot_visium_adipocyte_insulin_box_score_allc_new, fig.width=4, fig.height=3}
canno_merge <- canno_ins
canno_merge$cluster <- paste0("C", canno_ins$seurat_clusters)
c_keep <- subset(canno_merge, cluster_group %in% c("Adipocyte", "Fibroblast/Pread", "Vascular"))$cluster


d_plot <- cbind(log2(se_ins_data_score_allc_mean), my_sample_col_allc)
d_plot$order <- rep(seq(1:19), 4)
d_plot[d_plot$up==0,"up"] <- NA
d_plot[d_plot$down==0,"down"] <- NA
d_plot$up_score <- (d_plot$up - mean(d_plot$up, na.rm = T)) / sd(d_plot$up, na.rm = T)

d_plot <- subset(d_plot, cluster %in% c_keep)
d_plot2 <- merge(d_plot, canno_merge, by="cluster")
d_plot2$cluster <- as.factor(d_plot2$cluster)


p1 <- ggplot(d_plot2, aes(x=cluster_anno , y=up_score))+
  geom_hline(yintercept = 0, linetype="solid", color = "grey50") +
  geom_boxplot(width=0.5) +
  geom_point(size=1, aes(color=as.factor(m_value)), alpha=1) +
  coord_flip() +
  scale_x_discrete(limits = rev) +
  scale_color_manual(values = my_colour$m_value) +
  ylim(min(d_plot[,1:2])-.02, max(d_plot[,1:2])+0.02) +
  labs(y="Standardized log2-ratio") +
  labs(title="Insulin resp. gene signature", x="", color="M value") +
  theme_classic() +
  theme(legend.position = "bottom"); p1

p2 <- ggplot(d_plot2, aes(x=reorder(cluster_anno, up_score, na.rm = TRUE) , y=up_score))+
  geom_hline(yintercept = 0, linetype="solid", color = "grey50") +
  geom_boxplot(width=0.5) +
  geom_point(size=1, aes(color=as.factor(m_value)), alpha=1) +
  coord_flip() +
  scale_color_manual(values = my_colour$m_value) +
  ylim(min(d_plot[,1:2])-.02, max(d_plot[,1:2])+0.02) +
  labs(y="Standardized log2-ratio") +
  labs(title="Insulin resp. gene signature", x="", color="M value") +
  theme_classic() +
  theme(legend.position = "bottom"); p2

 
png_res <- 300
fname1 <- "bulk_dea_insulin_cage_visium-all_cluster_box_scores_selected-a"
png(file.path(DIR_FIG, paste0(fname1, ".png")), width = 4*png_res, height = 3*png_res, res = png_res);p1;dev.off()

fname2 <- "bulk_dea_insulin_cage_visium-all_cluster_box_scores_selected-b"
png(file.path(DIR_FIG, paste0(fname2, ".png")), width = 4*png_res, height = 3*png_res, res = png_res);p2;dev.off()


fname <- "bulk_dea_insulin_cage_visium-all_cluster_box_scores_selected.csv"
write.csv(d_plot2, file.path(DIR_RES, "tables", fname), quote = F, row.names = F)

```



<br><br>


# Wrap up
  
## Session information  
  
```{r sessioninfo}
Sys.Date()
sessionInfo()
```
