---
title: "Visium data - 'Insulin' samples: Cluster DEA before/after insulin stimulation"
author: "L. Franzén lovisa.franzen@scilifelab.se"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    theme: cosmo
    highlight: tango
    css: style.css
    code_folding: "hide"
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    number_sections: false
    
---

***
  
# Description  
  
Aim with this analysis is to look at insulin responses in all clusters, not only adipocytes. Two approaches are used: 1) Seurat's FindMarkers() function on spot level with normalized expression data, and 2) DESeq2 on pooled raw counts for each cluster and individual, where the design is set up to find DEGs before/after insulin stimulation with subject as covariate (~ subject + insulin).  
  
<br>
  
***

<br><br>  


# Initialize  

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = T,
	cache.lazy = FALSE
)

set.seed(42)

#' Load libs
library(Matrix)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(ggrepel)
library(cowplot)
library(patchwork)
library(ggpubr)

library(Seurat)
library(STutility)

library(DESeq2)

library(DT)
library(knitr)


#' Define analysis tag
ANALYSIS_ID <- "visium_insulin"


#' Define project paths
PROJECT_ID <- "visium"
DIR_WD <- getwd()
DIR_ROOT <- file.path(getwd(), "..")  # 

DIR_DATA <- file.path(DIR_ROOT, "data", PROJECT_ID)
DIR_RES <- file.path(DIR_ROOT, "results" , PROJECT_ID)
DIR_FIG <- file.path(DIR_RES, "figures")


#' Colors
source(file.path(DIR_WD, "colors.R"))

#' Define custom functions
'%!in%' <- function(x,y)!('%in%'(x,y));
```


<br><br> 

## Load necessary files and tables  
  
  
### Sample insulin cluster annotation table  

```{r read_canno}
c_anno <- read.csv(file.path(DIR_RES, "tables", "visium_insulin.clustering_annotations.csv"), stringsAsFactors = F)
```
<br><br>
  
### Metadata  

```{r read_metadata}
metadata <- read.table(file.path(DIR_DATA, "visium_sample_metadata.tsv"), sep = "\t", header = T, stringsAsFactors = F)
metadata$bmi <- as.numeric(metadata$bmi)
rownames(metadata) <- metadata$sample_name

metadata_ins <- subset(metadata, insulin_stim == 0 | insulin_stim == 1)
metadata_ins$subject_id_insulin <- paste0(metadata_ins$subject_id, "_", metadata_ins$insulin_stim)

datatable(metadata_ins)
```

  
<br><br>

### Seurat object  

```{r read_seobj}
se <- readRDS(file.path(DIR_RES, "se-object.visium_insulin.rds"))
se <- AddMetaData(se, paste0(se$subject_id, "_", se$insulin_stim), col.name = "subject_id_insulin")
```

<br><br>


# Spot data (Seurat)
  
## FindMarkers  

```{r clustercond_findmarkers}
se$cluster_cond <- paste(se$seurat_clusters, se$insulin_stim, sep = "_")
Idents(se) <- "cluster_cond"

clustercond_markers <- list()
for (c in seq_along(unique(se$seurat_clusters))) {
  if( paste0(c, "_1") %in% unique(se$cluster_cond) & paste0(c, "_0") %in% unique(se$cluster_cond) ){
      message("Finding markers between conditions within cluster ", c)
      cid <- paste0("C", c)
      clustercond_markers[[cid]] <- FindMarkers(se,
                                               ident.1 = paste0(c, "_1"), ident.2 = paste0(c, "_0"),
                                               only.pos = F,
                                               min.pct = 0.1,
                                               logfc.threshold = 0.15,
                                               verbose = T)
      clustercond_markers[[cid]]$gene <- rownames(clustercond_markers[[cid]])
      clustercond_markers[[cid]]$cluster <- cid
  } else {
    message(paste("Cluster", c, "is not available for all conditions --- excluded!"))
  }
}
```


```{r clustercond_findallmarkers}
clustercond_markers_all_df <- do.call(rbind.data.frame, clustercond_markers)
clustercond_markers_df <- clustercond_markers_all_df[clustercond_markers_all_df$p_val_adj < 0.05, ]
clustercond_markers_df <- clustercond_markers_df[, colnames(clustercond_markers_df)[c(7, 6, 1:5)]]

datatable(clustercond_markers_df, rownames = F)
```
<br><br>

  
## Plot  
  
Expression correlation  

```{r clustercond_findallmarkers_plot_scatter, fig.width=14, fig.height=18}
plist <- list()
for ( c in as.character(seq(1:19)) ){
  cluster_c <- subset(se, seurat_clusters == c)
  Idents(cluster_c) <- "insulin_stim"
  avg.cluster_c <- log1p(AverageExpression(cluster_c, verbose = FALSE)$RNA)  # natural log of avg expression
  avg.cluster_c$gene <- rownames(avg.cluster_c)
  
  label_genes <- clustercond_markers_df[clustercond_markers_df$cluster == paste0("C", c) & abs(clustercond_markers_df$avg_logFC) > 0.25, "gene"]
  
  p <- ggplot(avg.cluster_c, aes(`0`, `1`)) + 
          geom_point(alpha = .25, color = color_high) + 
          geom_abline(intercept = 0, slope = 1, color = "black") +
          scale_y_log10() + 
          scale_x_log10() +
          labs(x="Before insulin", y="After insulin", title = paste("Cluster", c)) +
          theme_classic() + theme(plot.title = element_text(hjust=0.5))
  
  if (length(label_genes)>0) {
    p <- p + geom_point(data = avg.cluster_c[label_genes, ], mapping = aes(`0`, `1`), alpha = .25, color = color_low)
    p <- LabelPoints(plot = p, points = label_genes, repel = TRUE, color = color_low)
  }
  plist[[c]] <- p
}

pgrid <- plot_grid(plotlist = plist, nrow = 5);pgrid
```
<br><br>

  
**Violin**  
```{r clustercond_findallmarkers_plot_violin, fig.width=18, fig.height=4}
p_vln <- VlnPlot(se, features = c("THRSP", "ANGPTL8", "PDK4"), 
                 split.by = "insulin_stim", group.by = "cluster_anno", 
                 cols = colors_insulin,
                 pt.size = 0, combine = T) &
            stat_summary(fun.y = median, geom = 'crossbar', colour = "black", width = 0.5); p_vln
```
<br><br>

  
**Vulcano plot**  
```{r clustercond_findallmarkers_plot_vulcano, fig.width=10, fig.height=12}
d_plot <- clustercond_markers_df
colnames(d_plot)[colnames(d_plot)=="p_val_adj"] <- "padj"

d_plot$sign_1m6 <- ifelse(test = d_plot$padj<1e-06 & !is.na(d_plot$padj), d_plot$gene, "")
d_plot$sign_001 <- ifelse(test = d_plot$padj<0.01 & !is.na(d_plot$padj), d_plot$gene, "")
d_plot <- subset(d_plot, !-log10(padj)<0.01)

d_plot$cluster_new <- factor(d_plot$cluster, levels = unique(d_plot$cluster))

p1 <- ggplot(d_plot, aes(x=avg_logFC, y=-log10(padj))) + 
  geom_vline(xintercept = 0, color = "grey50") +
  geom_hline(yintercept = 0, color = "black") +
  geom_point(color = "grey30") +
  geom_point(data = subset(d_plot,padj<1e-06), 
             mapping = aes(x=avg_logFC, y=-log10(padj)), 
             color = color_high, size=1) +
  geom_text_repel(data = subset(d_plot,padj<1e-06), 
                  mapping = aes(x=avg_logFC, y=-log10(padj), label = sign_1m6), size=3) +
  xlim(-1.15, 1.15) +
  facet_wrap(~ cluster_new, ncol=4) +
  labs(title = "Significance level of padj < 1e-06") +
  theme_linedraw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank());p1

p2 <- ggplot(d_plot, aes(x=avg_logFC, y=-log10(padj))) + 
  geom_vline(xintercept = 0, color = "grey50") +
  geom_hline(yintercept = 0, color = "black") +
  geom_point(color = "grey30") +
  geom_point(data = subset(d_plot, padj<0.01), 
             mapping = aes(x=avg_logFC, y=-log10(padj)), 
             color = color_high, size=1) +
  geom_text_repel(data = subset(d_plot, padj<0.01 & padj>1e-06), 
                  mapping = aes(x=avg_logFC, y=-log10(padj), label = sign_001), size=3) +
  xlim(-1.15, 1.15) +
  labs(title = "Significance level of padj < 0.001") +
  facet_wrap(~ cluster_new, ncol=4) +
  theme_linedraw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank());p2
```


<br><br>

## Export

```{r export_table_plot}
fname1 <- paste0(ANALYSIS_ID, ".clustercond_findmarkers_res")
pdf(file = file.path(DIR_RES, "figures", paste0(fname1, ".pdf")), width = 14, height = 18, useDingbats = F);pgrid;dev.off()
pdf(file = file.path(DIR_RES, "figures", paste0(fname1, "_violin", ".pdf")), width = 18, height = 4, useDingbats = F);p_vln;dev.off()
pdf(file = file.path(DIR_RES, "figures", paste0(fname1, "_vulcano", ".pdf")), width = 10, height = 12, useDingbats = F);p1;p2;dev.off()
write.csv(x = clustercond_markers_df, file = file.path(DIR_RES, "tables", paste0(fname1, ".csv")), row.names = F)
write.csv(x = clustercond_markers_all_df, file = file.path(DIR_RES, "tables", paste0(fname1, "_all.csv")), row.names = F)
```

<br><br>
  
  
# DESeq2 on pseudo-bulk

## Pool data
```{r bulk_data}
bulk_data_ins_list <- list()
bulk_data_ins <- data.frame(row.names = rownames(se))
s_names <- unique(se$sample_name)
c_include <- as.character(unique(se$seurat_clusters))

for (c in c_include) {
  for (s in s_names ) {
    
    n_spots <- nrow(subset(se[[]], sample_name == s & seurat_clusters == c))
    if (n_spots > 0) {
      s_spots <- colnames(subset(se, sample_name == s & seurat_clusters == c))
      st_data_s <- SubsetSTData(se, spots = s_spots)
      st_data_s_mat <- st_data_s@assays$RNA@counts
      
      if (n_spots > 1) {
        s_data <- data.frame(Matrix::rowSums(st_data_s_mat))
      } else {
        s_data <- data.frame(st_data_s_mat)
      }
      
    } else {
      s_data <- data.frame(c_s = rep(0, nrow(se)), row.names = rownames(se))
    }
    colnames(s_data) <- paste0(c, ".", s)
    bulk_data_ins <- cbind(bulk_data_ins, s_data)
  }
}

head(bulk_data_ins)
```
<br><br>
  
  
## Design table  

```{r de_design}
cols_keep <- c("insulin_stim", "tissue_id", "condition", "gender", "date_exp", "subject_id", "subject_alias", "subject_id_insulin")
de_des_ins <- metadata_ins[, cols_keep]
de_des_ins$date_exp <- gsub(pattern = "-", replacement = ".", x = de_des_ins$date_exp)
de_des_ins$insulin_stim <- as.factor(de_des_ins$insulin_stim)
de_des_ins$tissue <- unlist( lapply( strsplit(x = de_des_ins$tissue_id, split = "-"), function(x) head(x, n=1) ) )
de_des_ins$tissue_id <- NULL

de_des_ins <- de_des_ins[2:dim(de_des_ins)[1],]  # remove NK49-before from first batch
de_des_ins

```
<br><br>
  
  
## DESeq2

```{r deseq2_cluster_sample, message=TRUE}
c_include <- as.character(sort(unique(se$seurat_clusters)))
res_ins_clusters <- list()
for(c in c_include) {
  message(paste("DESeq2 for cluster", c))
  de_des_ins_cluster <- de_des_ins
  de_des_ins_cluster$ind.n <- as.factor((c(1, 1, 2, 2, 3, 3, 4, 4)))
  
  rownames(de_des_ins_cluster) <- paste0(c, ".", rownames(de_des_ins_cluster))
  
  s_ins <- rownames(de_des_ins_cluster)
  s_include <- s_ins[colSums(bulk_data_ins[, s_ins]) > 0]  # Remove samples with 0 counts in all genes
  
  d_ins <- bulk_data_ins[, colnames(bulk_data_ins) %in% s_include]
  de_des_ins_cluster_filt <- de_des_ins_cluster[colnames(d_ins), ]
  
  dds_ins_c <- DESeqDataSetFromMatrix(countData = d_ins,
                                      colData = de_des_ins_cluster_filt,
                                      design = ~ tissue + insulin_stim)
  dds_ins_c <- DESeq(dds_ins_c)

  res_ins_c <- results(dds_ins_c, name="insulin_stim_1_vs_0")
  res_ins_c_df <- as.data.frame(res_ins_c)

  res_ins_clusters[[c]] <- res_ins_c_df
  message("======================")
}
```


```{r deseq2_cluster_sample_significant}
res_ins_clusters_sign <- list()
res_ins_clusters_all <- list()
for(c in names(res_ins_clusters)) {
  d_tmp <- res_ins_clusters[[c]]
  d_tmp$cluster <- paste0("C", c)
  d_tmp$gene <- rownames(d_tmp)
  d_tmp <- d_tmp[, colnames(d_tmp)[c(7, 8, 1:6)]]
  
  res_ins_clusters_all[[c]] <- d_tmp
  
  d_tmp2 <- subset(d_tmp, padj < 0.05)
  if (nrow(d_tmp2) > 0) {
    res_ins_clusters_sign[[c]] <- d_tmp2
  }
}

res_ins_clusters_sign_df <- do.call(rbind.data.frame, res_ins_clusters_sign)
res_ins_clusters_all_df <- do.call(rbind.data.frame, res_ins_clusters_all)

datatable(res_ins_clusters_sign_df, rownames = F)
```

<br><br>


## Plot

```{r deseq2_plot_vulcano, fig.width=10, fig.height=11}
d_plot <- res_ins_clusters_all_df
d_plot <- d_plot[d_plot$cluster %in% unique(res_ins_clusters_sign_df$cluster), ]
d_plot$sign_001 <- ifelse(test = d_plot$padj<0.01 & !is.na(d_plot$padj), d_plot$gene, "")
d_plot$sign_05 <- ifelse(test = d_plot$padj<0.05 & !is.na(d_plot$padj), d_plot$gene, "")
d_plot <- subset(d_plot, !-log10(padj)<0.01)

d_plot$cluster_new <- factor(d_plot$cluster, levels = unique(d_plot$cluster))

p1 <- ggplot(d_plot, aes(x=log2FoldChange, y=-log10(padj))) + 
  geom_vline(xintercept = 0, color = "grey50") +
  geom_hline(yintercept = 0, color = "black") +
  geom_point(color = "grey30") +
  geom_point(data = subset(d_plot,padj<0.01), 
             mapping = aes(x=log2FoldChange, y=-log10(padj)), 
             color = color_high, size=1) +
  geom_text_repel(data = subset(d_plot,padj<0.01), 
                  mapping = aes(x=log2FoldChange, y=-log10(padj), label = sign_001), size=3) +
  xlim(-6, 6) +
  facet_wrap(~ cluster_new, ncol=4) +
  labs(title = "Significance level of padj < 0.01") +
  theme_linedraw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank());p1

p2 <- ggplot(d_plot, aes(x=log2FoldChange, y=-log10(padj))) + 
  geom_vline(xintercept = 0, color = "grey50") +
  geom_hline(yintercept = 0, color = "black") +
  geom_point(color = "grey30") +
  geom_point(data = subset(d_plot, padj<0.05), 
             mapping = aes(x=log2FoldChange, y=-log10(padj)), 
             color = color_high, size=1) +
  geom_text_repel(data = subset(d_plot, padj<0.05 & padj>0.01), 
                  mapping = aes(x=log2FoldChange, y=-log10(padj), label = sign_05), size=3) +
  xlim(-6, 6) +
  facet_wrap(~ cluster_new, ncol=4) +
  labs(title = "Significance level of padj < 0.05") +
  theme_linedraw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank());p2
```

<br><br>


## Export  

```{r deseq2_export}
fname2 <- paste0(ANALYSIS_ID, ".clustercond_deseq2_res")
pdf(file = file.path(DIR_RES, "figures", paste0(fname2, "_vulcano.pdf")), width = 10, height = 11, useDingbats = F);p1;p2;dev.off()
# pdf(file = file.path(DIR_RES, "figures", paste0(fname2, "_sign001.pdf")), width = 10, height = 10, useDingbats = F);p1;dev.off()
# pdf(file = file.path(DIR_RES, "figures", paste0(fname2, "_sign05.pdf")), width = 10, height = 10, useDingbats = F);p2;dev.off()
write.csv(x = res_ins_clusters_sign_df, file = file.path(DIR_RES, "tables", paste0(fname2, ".csv")), row.names = F)
write.csv(x = res_ins_clusters_all_df, file = file.path(DIR_RES, "tables", paste0(fname2, "_all.csv")), row.names = F)
```

<br><br>


# Gene expresssion per subject

```{r compare_subject_gene_expr}
genes_plot <- subset(res_ins_clusters_sign_df, cluster == "C4" & log2FoldChange > 2)$gene

gene_bulk_df <- t(bulk_data_ins[genes_plot, ])

c_include <- as.character(sort(as.numeric( unique(unlist(lapply(strsplit(x = rownames(gene_bulk_df), split = "\\."), `[[`, 1))) )))
subject_info <- de_des_ins
subject_info$sample_name <- rownames(de_des_ins)

for(c in c_include){
  subject_info_c <- subject_info
  subject_info_c$cluster <- c
  rownames(subject_info_c) <- paste0(c, ".", subject_info_c$sample_name)
  
  if(c == "1"){
    subject_info_all <- subject_info_c
  } else {
    subject_info_all <- rbind(subject_info_all, subject_info_c)
  }
}

r_include <- intersect(rownames(gene_bulk_df), rownames(subject_info_all))
gene_bulk_df2 <- cbind(gene_bulk_df[r_include, ], subject_info_all[r_include, ])
gene_bulk_df2$cluster_new <- factor(gene_bulk_df2$cluster, levels = as.character(sort(as.numeric(unique(gene_bulk_df2$cluster)))) )
gene_bulk_df2[, genes_plot][gene_bulk_df2[ , genes_plot]<1] <- NA
```


## Plot  

```{r export_fname3}
fname3 <- paste0(ANALYSIS_ID, ".clustercond_subject_expression")
```


```{r compare_subject_gene_expr_plot_bar, fig.width=10, fig.height=10}
pdf(file = file.path(DIR_RES, "figures", paste0(fname3, "_pooledcounts.pdf")), width = 10, height = 10, useDingbats = F)
for(g in genes_plot) {
  p <- ggplot(gene_bulk_df2, aes_string(x="tissue", y=g, fill = "insulin_stim")) +  # subset(gene_bulk_df2, cluster=="1")
    geom_col(position = "dodge") +
    scale_fill_manual(values = colors_insulin) +
    facet_wrap(~cluster_new, ncol = 5) +  # , scales = "free_y"
    labs(x="", y="Pooled count", title=g, fill ="") +
    scale_y_log10() +
    annotation_logticks(size = .2, sides = "l") +
    theme_linedraw() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid.major = element_line(color="grey70"),
          panel.grid.minor = element_blank())
  print(p)
}
dev.off()
```


```{r compare_subject_gene_expr_plot_violin, fig.height=14, fig.width=14, warning=FALSE}
genes_plot2 <- intersect(rownames(se@assays$SCT@scale.data), genes_plot)

pdf(file = file.path(DIR_RES, "figures", paste0(fname3, "_spots.pdf")), width = 14, height = 14, useDingbats = F)
for(s in sort(unique(se$subject_id))){
  p_vln <- VlnPlot(subset(se, subject_id == s), features = genes_plot2, 
                   group.by = "seurat_clusters", split.by = "insulin_stim", 
                   slot = "scale.data", 
                   pt.size = 0.0, cols = colors_insulin,
                   ncol = 2) & geom_hline(yintercept = 0) & labs(x=paste(s, "clusters"), y="Scaled expr.")
  print(p_vln) 
}
dev.off()
```



<br><br>


# Integration of data

Update: This approach is perhaps not quite suitable. --- skip


## Split and perform integration  

https://satijalab.org/seurat/v3.0/immune_alignment.html
  
  
1. Split data

```{r split_data}
# se <- AddMetaData(se, paste0(se$subject_id, "_", se$insulin_stim), col.name = "subject_id_insulin")
# 
# spots_ins0 <- colnames(subset(se, insulin_stim == 0))
# spots_ins1 <- colnames(subset(se, insulin_stim == 1))
# 
# se_ins0 <- SubsetSTData(se, spots = spots_ins0)
# se_ins1 <- SubsetSTData(se, spots = spots_ins1)
```
  
  
2. Perform integration  

```{r integrate_data_P1}
# se_anchors <- FindIntegrationAnchors(object.list = list(se_ins0, se_ins1), reduction = "cca", dims = 1:30)
```

```{r integrate_data_P2}
# se_combined <- IntegrateData(anchorset = se_anchors, dims = 1:30, features.to.integrate = rownames(se@assays$SCT@data))
# DefaultAssay(se_combined) <- "integrated"
```

```{r scale_data}
# se_combined <- ScaleData(se_combined)
```
  
  
3. Check integration

```{r plot_integrated_data, fig.width=16, fig.height=4}
# VlnPlot(se_combined, features = c("PLIN4", "ADIPOQ", "LEP"), 
#         group.by = "seurat_clusters", split.by = "insulin_stim", 
#         assay = "integrated", slot = "data", 
#         pt.size = 0)
# 
# VlnPlot(se_combined, features = c("THRSP", "LPL", "BHLHE40"), 
#         split.by = "date_exp", 
#         group.by = "tissue_id", 
#         assay = "SCT", slot = "data", 
#         pt.size = 0)
```


```{r find_markers}
# markers_all <- FindAllMarkers(se_combined, )
```




***  

# Wrap up

<br><br>

## Session information  
  
This analysis was last compiled on `r Sys.Date()`.  
  
<br>
  
```{r sessioninfo}
sessionInfo()
```

